---
title: "Executive functions data preparation"
author: "Rebecca Delfendahl and Charlotte Meinke"
date: "2023-10-05"
output: html_document
---

```{r R Setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,rows.print = 6, cols.min.print = 3)

library(haven)
library(rmatio)
library(DescTools)
library(data.table)
library(tidyverse)
library(caret)
library(ggplot2)
```

# 0 Initial configurations

## Define functions

```{r Define functions}

########################################
## Function 1: get relevant data from mat-file for a single subject
########################################
df_from_matlab <- function(task_name, subject){
  file_name = paste("SFB_C5_",subject,"_",task_name,".mat", sep = "")
  data_path = file.path(basic_path,file_name)
  data <- read.mat(data_path)
  # So far, we only need TrialList_Log_ -> it is the combination of TL_ und LogMat
  subtask = paste("TrialList_Log_",task_name,sep = "")
  task_df <- as.data.frame(rbindlist(data[[subtask]]))
  header <- paste(subtask, "Header", sep = "_")
  colnames(task_df) <- c(rbindlist(data[[header]])[1,])
  return(task_df)
}

########################################
## Function 2: merge all task-data of all subjects in one dataframe
########################################
combine_df_across_subj <- function(task){
  for (subject in subjects){
    try({
    df <- df_from_matlab(task,subject)
    df$subject <- subject
    df_all_subj <- rbind(df_all_subj,df)
})
  }
  return(df_all_subj)
}

########################################
## Function 3: preprocessing and calculation of mean RT and PC
########################################
preprocess_RT_PC <- function(task, subjects){
  
  # Initialize empty holder dataframe
  AllData <- data.frame()
  
  # Loop across all participants
  for (subject in subjects) {
    # Import data
    data <- df_from_matlab(task_name = task, subject = subject)
    
    # Bring data NumberLetter and data Stroop to same format
    if (task == "NumberLetter"){
      colnames(data) <- sub(" ", "_", colnames(data))
      data <- data %>%
        rename("Condition"="Task_transition")
    } else if (task == "Stroop"){
        data <- data %>%
          rename("Accuracy"="Fehler")
    }
    
    # Basic preprocessing
    data_red <- data %>%
      # Select relevant columns
      select(Condition, Response, Accuracy, RT) %>%
      # Discard first trial (only applies to NumbLet)
      filter(Condition != "NoPreviousTrial")
    
    # Convert RTs and Accuracy to numeric
    data_red$RT <- as.numeric(data_red$RT)
    data_red$Accuracy <- as.numeric(data_red$Accuracy)
    
    # STEP 1.1.: Processing data for RT analysis
    
    relevantData_RT <- data_red %>%
      ## Step 1: Remove wrong responses
      filter(Accuracy == 1) %>%
      ## Step 2: Change RT to ms
      mutate(RT = RT * 1000)
    
    relevantData_RT_clean <- relevantData_RT %>%
      ## Step 3: Remove outlier 
      filter(RT > 150)
    
    # STEP 1.2.: Compute mean RT overall and per condition 
    
    overall_RT_mean <- mean(relevantData_RT_clean$RT)
    
    RT_means <- relevantData_RT_clean %>%
      group_by(Condition) %>%
      summarize_at(vars(RT), mean)
    
    ## Transform data from LONG to WIDE format
    wideData_RT <- RT_means %>%
      pivot_wider(names_from = Condition, values_from = RT)
    
      
    # STEP 2: Compute proportion correct overall and per condition 
    
    overall_PC <- mean(data_red$Accuracy)
    
    PC <- data_red %>%
      group_by(Condition) %>%
      summarize_at(vars(Accuracy), mean)
    
    ## TODO: Convert to percentages?
    
    ## Transform data from LONG to WIDE format
    wideData_PC <- PC %>%
      pivot_wider(names_from = Condition, values_from = Accuracy)
    
    # STEP 3: Merging the two wide dataframe
    
    ## Update column names to include the variable identity
    colnames(wideData_RT) <- paste(colnames(wideData_RT),"RT", sep="_")
    colnames(wideData_PC) <- paste(colnames(wideData_PC),"PC", sep="_")
    
    ## Merge dataframes
    wideData_merged <- cbind(wideData_RT, wideData_PC)
    
    # STEP 4: Adding extra relevant info
    
    ## Compute % of removed trials
    percentRmv_RT <- 100 * (1 - (nrow(relevantData_RT_clean) / nrow(relevantData_RT)))
    wideData_merged$PercentRmv_RT <- percentRmv_RT
    
    ## Add subject ID, overall mean RT and overall PC
    wideData_merged$Subject <- subject
    wideData_merged$Overall_RT <- overall_RT_mean
    wideData_merged$Overall_PC <- overall_PC
    
    ## Store the current processed data into holder as a new row
    AllData <- rbind(AllData, wideData_merged)
  }
  return(AllData)
}

# Function 4: calculate mean and SD of RT and PC of healthy subjects
# mean_sd_HC <- function(data_HC){
#   # Calculate mean and sd of RT and PC across all healthy subjects and conditions
#   meanRT_HC <- mean(data_HC$Overall_RT)
#   sdRT_HC <- sd(data_HC$Overall_RT)
#   meanPC_HC <- mean(data_HC$Overall_PC)
#   sdPC_HC <- sd(data_HC$Overall_PC)
#   return(meanRT_HC, sdRT_HC, meanPC_HC, sdPC_HC) # TODO: have to be stored in a list
# }

```

## Load data

```{r Load data}
basic_path = "Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Rohdaten/Task_battery/Task_battery"

# Load data
data_all <- read_dta('Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Rohdaten/CRC_C5_Hilbert/CRC_C5_Hilbert/Data_Kevin_28.07.23.dta')
data_tasks_old <- read_dta('Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Rohdaten/CRC_C5_Hilbert/CRC_C5_Hilbert/TaskBattery.dta')

```

## Reduce data_all 
Keep only ID, group identity, FAS_T1 and FAS_T3 (response criterion)

```{r Reduce df data_all}
  
data_all_red <- data_all %>%
  rename(Subject = id) %>%
  select(Subject, Gruppe, T1_BAT_FAS_score, T3_BAT_FAS_score)

# Rename stid for later merge
data_tasks_old <- data_tasks_old %>%
  rename(Subject = stid)
```

# 1. Import Data

For each subject, we have 5 matlab dataframes:
- TwoBack
- StopSignal
- Stroop
- Stroop_mouse
- NumberLetter
- ExpOrder

## Get list of all subjects

```{r List of all subjects}

files <- list.files(basic_path)

sub_first <- gsub("SFB_C5_","",files)
sub_IDs <- sapply(strsplit(sub_first, "_"), function(x) x[1])
subjects <- unique(sub_IDs)
# Delete alt and "" from subjects
subjects <- subjects[subjects != "alt"]
subjects <- subjects[subjects != ""]
```

## Optional (for visualization): Import data from single subject
```{r import_data}
data_twoback <- df_from_matlab("TwoBack")
data_stop <- df_from_matlab("StopSignal")
data_stroop <- df_from_matlab("Stroop")
data_nl <- df_from_matlab("NumberLetter")
# TODO: Stroop mouse data has another data structure, so we need an extra function/ modification (Exclude history lists (more than 160 variables))
#df_stroop_mouse <- rbindlist(data_stroop_mouse[["triallog"]])[1,]
```

## Optional: Combine all subjects and all tasks in one dataframe
```{r combine_data}

task = "TwoBack"
tasks = c("TwoBack","StopSignal","Stroop","NumberLetter")
df_all_subj = data.frame()

#subjects = subjects[3:4]
all_TwoBack <- combine_df_across_subj(task = "TwoBack")
all_Stroop <- combine_df_across_subj(task = "Stroop")
all_StopSignal <-  combine_df_across_subj(task = "StopSignal")
all_NumberLetter <- combine_df_across_subj(task = "NumberLetter")

```

# 2. Preprocess data of each subject for Number-Letter, Stroop and Two-Back task
```{r Looping across all participants, warning=FALSE}

AllData_NumbLet <- preprocess_RT_PC(task = "NumberLetter", subjects = subjects)
AllData_Stroop <- preprocess_RT_PC(task = "Stroop", subjects = subjects)
AllData_TwoBack <- preprocess_RT_PC(task = "TwoBack", subjects = subjects)
```

## Optional: save the processed data
```{r Save preprocessed data as .csv}

write.csv(AllData_NumbLet, "Z:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Dateien_Github/Tasks_Merged Data/AllData_NumbLet.csv")
write.csv(AllData_Stroop, "Z:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Dateien_Github/Tasks_Merged Data/AllData_Stroop.csv")
write.csv(AllData_TwoBack, "Z:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Dateien_Github/Tasks_Merged Data/AllData_TwoBack.csv")
```

# 3. Merge task data with data_all
```{r Merge task data with data_all_red}
task_dfs <- list(AllData_NumbLet, AllData_Stroop, AllData_TwoBack)
result_dfs <- list()

for (task in task_dfs){
  # Step 1: convert Subject to numeric for use in left_join
  task$Subject <- as.numeric(task$Subject)
  # Step 2: merge task dataframes and data_all_red
  result_dfs[[length(result_dfs) + 1]] <- left_join(
    x = task,
    y = data_all_red,
    by = "Subject")
}

AllData_NumbLet_merged <- result_dfs[[1]]
AllData_Stroop_merged <- result_dfs[[2]]
AllData_TwoBack_merged <- result_dfs[[3]]

```

# 4. Recode patient IDs from post-measurement

```{r Post IDs}
# NumberLetter
AllData_NumbLet_merged <- AllData_NumbLet_merged %>%
  mutate(Gruppe = ifelse(Subject >= 216601 & Subject <= 216760, 2, Gruppe))
# Stroop
AllData_Stroop_merged <- AllData_Stroop_merged %>%
  mutate(Gruppe = ifelse(Subject >= 216601 & Subject <= 216760, 2, Gruppe))
# TwoBack
AllData_TwoBack_merged <- AllData_TwoBack_merged %>%
  mutate(Gruppe = ifelse(Subject >= 216601 & Subject <= 216760, 2, Gruppe))
```

# 5. Exclusion of subjects for each task individually, separate HC from patients
Exclude subjects with missing group information and patients with missing FAS_T1 and/or FAS_T3. No subjects have NAs in all 4 task variables, therefore this exclusion step is not included anymore.

## Number-Letter Task
```{r Subject exclusion NumberLetter}

# Remove subjects with missing group information 
AllData_NumbLet_clean.1 <- AllData_NumbLet_merged %>%
  filter(!is.na(Gruppe))
excl_subs_NL_group <- anti_join(AllData_NumbLet_merged, AllData_NumbLet_clean.1, by = "Subject")

# Remove patients with missing FAS_T1 and/or FAS_T3
AllData_NumbLet_clean.2 <- AllData_NumbLet_clean.1 %>%
  filter(Gruppe == 0 | Gruppe == 2 | (Gruppe == 1 & !is.na(T1_BAT_FAS_score) & !is.na(T3_BAT_FAS_score)))
excl_subs_NL_rc <- anti_join(AllData_NumbLet_clean.1, AllData_NumbLet_clean.2, by = "Subject")

# Remove subjects that have NAs in all 4 task variables
# AllData_NumbLet_clean.3 <- AllData_NumbLet_clean.2 %>%
#   filter(!all(is.na(Repeat_RT), is.na(Switch_RT), is.na(Repeat_PC), is.na(Switch_PC)))
# excl_subs_NL_task <- anti_join(AllData_NumbLet_clean.2, AllData_NumbLet_clean.3, by = "Subject")
```
`r nrow(excl_subs_NL_group)` subjects were removed due to missing group information 
`r nrow(excl_subs_NL_rc)` patients were removed due to missing FAS_T1 and/or FAS_T3 (i.e. response criterion)

## Stroop Task
```{r Subject exclusion Stroop}

# Remove subjects with missing group information 
AllData_Stroop_clean.1 <- AllData_Stroop_merged %>%
  filter(!is.na(Gruppe))
excl_subs_Str_group <- anti_join(AllData_Stroop_merged, AllData_Stroop_clean.1, by = "Subject")

# Remove patients with missing FAS_T1 and/or FAS_T3
AllData_Stroop_clean.2 <- AllData_Stroop_clean.1 %>%
  filter(Gruppe == 0 | Gruppe == 2 | (Gruppe == 1 & !is.na(T1_BAT_FAS_score) & !is.na(T3_BAT_FAS_score)))
excl_subs_Str_rc <- anti_join(AllData_Stroop_clean.1, AllData_Stroop_clean.2, by = "Subject")

# Remove subjects that have NAs in all 4 task variables
# AllData_Stroop_clean.3 <- AllData_Stroop_clean.2 %>%
#   filter(!all(is.na(Congruent_RT), is.na(Incongruent_RT), is.na(Congruent_PC), is.na(Incongruent_PC)))
# excl_subs_Str_task <- anti_join(AllData_Stroop_clean.2, AllData_Stroop_clean.3, by = "Subject")
```
`r nrow(excl_subs_Str_group)` subjects were removed due to missing group information 
`r nrow(excl_subs_Str_rc)` patients were removed due to missing FAS_T1 and/or FAS_T3 (i.e. response criterion)

## TwoBack Task
```{r Subject exclusion TwoBack}

# Remove subjects with missing group information 
AllData_TwoBack_clean.1 <- AllData_TwoBack_merged %>%
  filter(!is.na(Gruppe))
excl_subs_2B_group <- anti_join(AllData_TwoBack_merged, AllData_TwoBack_clean.1, by = "Subject")

# Remove patients with missing FAS_T1 and/or FAS_T3
AllData_TwoBack_clean.2 <- AllData_TwoBack_clean.1 %>%
  filter(Gruppe == 0 | Gruppe == 2 | (Gruppe == 1 & !is.na(T1_BAT_FAS_score) & !is.na(T3_BAT_FAS_score)))
excl_subs_2B_rc <- anti_join(AllData_TwoBack_clean.1, AllData_TwoBack_clean.2, by = "Subject")

# Remove subjects that have NAs in all 4 task variables
# AllData_TwoBack_clean.3 <- AllData_TwoBack_clean.2 %>%
#   filter(!all(is.na(Foil_RT), is.na(Target_RT), is.na(Foil_PC), is.na(Target_PC)))
# excl_subs_2B_task <- anti_join(AllData_TwoBack_clean.2, AllData_TwoBack_clean.3, by = "Subject")
```
`r nrow(excl_subs_2B_group)` subjects were removed due to missing group information 
`r nrow(excl_subs_2B_rc)` patients were removed due to missing FAS_T1 and/or FAS_T3 (i.e. response criterion)

## Summary subject exclusion
IDs of task-data with no group identity (8 in total):
- 1, 2: correspond to 216001 und 216002?
- 2160031: one 0 too much? (216031 exists)
- 216504: ???
- 2166108, 2166120, 2166121, 2166321: MZP2 and one 6 too much?

For each task, 19 patients had to be removed due to missing data for FAS_T1 and/or FAS_T3


# TO BE MODIFIED... (write one or two functions to use for all three tasks)

# 6. Calculate BIS (and Difference Score) for each condition and subject separately
BIS = Z(PC) - Z(RT)

Difference Switch and Repeat = BIS(Switch) - BIS(Repeat)
Difference Incongruent and Congruent = BIS(Congruent) - BIS(Incongruent)

## Number-Letter Task

```{r Scores Number-Letter Task}

## Step 1: Calculate mean and SD for RT and PC across all healthy subjects and all conditions
mean_sd_HC_NumbLet <- AllData_NumbLet_clean.2 %>% 
  filter(Gruppe == 0) %>%
  select(Overall_RT, Overall_PC) %>%
  summarise(meanRT = mean(Overall_RT),
            meanPC = mean(Overall_PC),
            sdRT = sd(Overall_RT),
            sdPC = sd(Overall_PC))

## Step 2: Standardize RT and PC for each condition
AllData_NumbLet_clean.2 <- AllData_NumbLet_clean.2 %>%
  mutate(Z_Repeat_RT = (Repeat_RT - mean_sd_HC_NumbLet$meanRT) / mean_sd_HC_NumbLet$sdRT,
         Z_Switch_RT = (Switch_RT - mean_sd_HC_NumbLet$meanRT) / mean_sd_HC_NumbLet$sdRT,
         Z_Repeat_PC = (Repeat_PC - mean_sd_HC_NumbLet$meanPC) / mean_sd_HC_NumbLet$sdPC,
         Z_Switch_PC = (Switch_PC - mean_sd_HC_NumbLet$meanPC) / mean_sd_HC_NumbLet$sdPC)

# Step 3: Calculate BIS(Repeat), BIS(Switch) and Difference Score
AllData_NumbLet_clean.2 <- AllData_NumbLet_clean.2 %>%
  mutate(BIS_Repeat = Z_Repeat_PC - Z_Repeat_RT,
         BIS_Switch = Z_Switch_PC - Z_Switch_RT) %>%
  mutate(Diff_Score_NumbLet = BIS_Switch - BIS_Repeat)

```

## Stroop Task

```{r Scores Stroop Task}

## Step 1: Calculate mean and SD for RT and PC across all healthy subjects and all conditions
mean_sd_HC_Stroop <- AllData_Stroop_clean.2 %>% 
  filter(Gruppe == 0) %>%
  select(Overall_RT, Overall_PC) %>%
  summarise(meanRT = mean(Overall_RT),
            meanPC = mean(Overall_PC),
            sdRT = sd(Overall_RT),
            sdPC = sd(Overall_PC))

## Step 2: Standardize RT and PC for each condition
AllData_Stroop_clean.2 <- AllData_Stroop_clean.2 %>%
  mutate(Z_Congruent_RT = (Congruent_RT - mean_sd_HC_Stroop$meanRT) / mean_sd_HC_Stroop$sdRT,
         Z_Incongruent_RT = (Incongruent_RT - mean_sd_HC_Stroop$meanRT) / mean_sd_HC_Stroop$sdRT,
         Z_Congruent_PC = (Congruent_PC - mean_sd_HC_Stroop$meanPC) / mean_sd_HC_Stroop$sdPC,
         Z_Incongruent_PC = (Incongruent_PC - mean_sd_HC_Stroop$meanPC) / mean_sd_HC_Stroop$sdPC)

# Step 3: Calculate BIS(Congruent), BIS(Incongruent) and Difference Score
AllData_Stroop_clean.2 <- AllData_Stroop_clean.2 %>%
  mutate(BIS_Congruent = Z_Congruent_PC - Z_Congruent_RT,
         BIS_Incongruent = Z_Incongruent_PC - Z_Incongruent_RT) %>%
  mutate(Diff_Score_Stroop = BIS_Incongruent - BIS_Congruent)
```

## TwoBack Task
```{r BIS TwoBack Task}

## Step 1: Calculate mean and SD for RT and PC across all healthy subjects and all conditions
mean_sd_HC_TwoBack <- AllData_TwoBack_clean.2 %>% 
  filter(Gruppe == 0) %>%
  select(Overall_RT, Overall_PC) %>%
  summarise(meanRT = mean(Overall_RT),
            meanPC = mean(Overall_PC),
            sdRT = sd(Overall_RT),
            sdPC = sd(Overall_PC))

## Step 2: Standardize RT and PC for each condition
AllData_TwoBack_clean.2 <- AllData_TwoBack_clean.2 %>%
  mutate(Z_Total_RT = (Overall_RT - mean_sd_HC_TwoBack$meanRT) / mean_sd_HC_TwoBack$sdRT,
         Z_Target_RT = (Target_RT - mean_sd_HC_TwoBack$meanRT) / mean_sd_HC_TwoBack$sdRT,
         Z_Foil_RT = (Foil_RT - mean_sd_HC_TwoBack$meanRT) / mean_sd_HC_TwoBack$sdRT,
         Z_Total_PC = (Overall_PC - mean_sd_HC_TwoBack$meanPC) / mean_sd_HC_TwoBack$sdPC,
         Z_Target_PC = (Target_PC - mean_sd_HC_TwoBack$meanPC) / mean_sd_HC_TwoBack$sdPC,
         Z_Foil_PC = (Foil_PC - mean_sd_HC_TwoBack$meanPC) / mean_sd_HC_TwoBack$sdPC)

# Calculate BIS(Target), BIS(Total) and BIS_Foil
AllData_TwoBack_clean.2 <- AllData_TwoBack_clean.2 %>%
  mutate(BIS_Total = Z_Total_PC - Z_Total_RT,
         BIS_Target = Z_Target_PC - Z_Target_RT,
         BIS_Foil = Z_Foil_PC - Z_Foil_RT)
```

# 7. Create separate dataframe for each group (HC, patient_pre, patient_post)

```{r Dataframe for each group}
# NumberLetter
split_df_NumbLet <- split(AllData_NumbLet_clean.2, AllData_NumbLet_clean.2$Gruppe)
data_NumbLet_HC <- split_df_NumbLet$"0"
data_NumbLet_Pat_pre <- split_df_NumbLet$"1"
data_NumbLet_Pat_post <- split_df_NumbLet$"2"

# Stroop
split_df_Stroop <- split(AllData_Stroop_clean.2, AllData_Stroop_clean.2$Gruppe)
data_Stroop_HC <- split_df_Stroop$"0"
data_Stroop_Pat_pre <- split_df_Stroop$"1"
data_Stroop_Pat_post <- split_df_Stroop$"2"

# TwoBack
split_df_TwoBack <- split(AllData_TwoBack_clean.2, AllData_TwoBack_clean.2$Gruppe)
data_TwoBack_HC <- split_df_TwoBack$"0"
data_TwoBack_Pat_pre <- split_df_TwoBack$"1"
data_TwoBack_Pat_post <- split_df_TwoBack$"2"
```

# 8. Save dataframes for t-tests
```{r Save dataframes}

# define function for ... purposes
write_data_to_csv <- function(data, filename) {
  base_path <- "Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/Without_wrong_responses/"
  write.csv(data, paste0(base_path, filename), row.names = FALSE)
}

# Write data to CSV using the function
write_data_to_csv(data_NumbLet_HC, "NumberLetter_HC.csv")
write_data_to_csv(data_NumbLet_Pat_pre, "NumberLetter_Patients_Pre.csv")
write_data_to_csv(data_NumbLet_Pat_post, "NumberLetter_Patients_Post.csv")

write_data_to_csv(data_Stroop_HC, "Stroop_HC.csv")
write_data_to_csv(data_Stroop_Pat_pre, "Stroop_Patients_Pre.csv")
write_data_to_csv(data_Stroop_Pat_post, "Stroop_Patients_Post.csv")

write_data_to_csv(data_TwoBack_HC, "TwoBack_HC.csv")
write_data_to_csv(data_TwoBack_Pat_pre, "TwoBack_Patients_Pre.csv")
write_data_to_csv(data_TwoBack_Pat_post, "TwoBack_Patients_Post.csv")
```


# 12. Distribution plots for Number-Letter Task

## Distribution of RT for each condition and group
```{r Number-Letter: Distribution of RT}

# Merge data from HC and patients, bring to long-format (to display in facet)
NumbLet_combined <- rbind(data_NumbLet_HC, data_NumbLet_Pat_pre[,1:12])
NumbLet_subset_RT <- NumbLet_combined[c("Repeat_RT", "Switch_RT", "Gruppe")]
NumbLet_subset_RT_long <- gather(NumbLet_subset_RT, key = "Condition", value = "RT", Repeat_RT, Switch_RT)

# Distribution plot of RT for both conditions and groups
plot_NumbLet_RT <- ggplot(NumbLet_subset_RT_long, aes(x=RT, fill=as.factor(Gruppe))) + 
  geom_histogram(binwidth=55, colour="black", fill="white") + 
  facet_grid(Gruppe ~ Condition) +
  labs(title = "RT Distribution Number-Letter Task by Group and Condition",
       x = "Reaction Time",
       y = "Frequency")
```

## HC: Distribution of mean RT across all conditions
```{r Number-Letter: Distribution of RT for only HCs}
mean_RT_HC = mean(data_NumbLet_HC$Overall_RT)

plot_NumbLet_RT_HC <- ggplot(data_NumbLet_HC, aes(x=Overall_RT)) +
  geom_histogram(binwidth=55, colour="black", fill="white") +
  geom_vline(xintercept = mean_RT_HC, 
             col = "red") +
  scale_x_continuous(breaks = seq(400, 2000, 400), lim = c(400, 2000)) +
  labs(title="Distribution of mean RT across conditions for HCs", 
       x="Reaction Time", 
       y="Frequency")
```

## Distribution of PC for each condition and group
```{r Number-Letter: Distribution of PC}

# Subset and bring to long-format
NumbLet_subset_PC <- NumbLet_combined[c("Repeat_PC", "Switch_PC", "Gruppe")]
NumbLet_subset_PC_long <- gather(NumbLet_subset_PC, key = "Condition", value = "PC", Repeat_PC, Switch_PC)

# Distribution plot of PC for both conditions and groups
plot_NumbLet_PC <- ggplot(NumbLet_subset_PC_long, aes(x=PC, fill=as.factor(Gruppe))) + 
  geom_histogram(binwidth=0.03, colour="black", fill="white") + 
  facet_grid(Gruppe ~ Condition) +
  labs(title = "PC Distribution Number-Letter Task by Group and Condition",
       x = "Proportion Correct",
       y = "Frequency")
```

## HC: Distribution of mean PC across all conditions
```{r Number-Letter: Distribution of PC for only HCs}
mean_PC_HC = mean(data_NumbLet_HC$Overall_PC)

plot_NumbLet_PC_HC <- ggplot(data_NumbLet_HC, aes(x=Overall_PC)) +
  geom_histogram(binwidth=0.03, colour="black", fill="white") +
  geom_vline(xintercept = mean_PC_HC, 
             col = "red") +
  scale_x_continuous(breaks = seq(0.4, 1.0, 0.2), lim = c(0.4, 1.1)) +
  labs(title="Distribution of mean PC across conditions for HCs", 
       x="Proportion Correct", 
       y="Frequency")
```

## Distribution of BIS for each condition 
```{r Number-Letter: Distribution of BIS & Diff_Score}

# Bring to long-format (to display in facet)
NumbLet_subset_BIS <- data_NumbLet_Pat_pre[c("BIS_Repeat", "BIS_Switch", "Diff_Score")]
NumbLet_subset_BIS_long <- gather(NumbLet_subset_BIS, key = "Score", value = "BIS", BIS_Repeat, BIS_Switch, Diff_Score)

# Distribution plot of BIS for both conditions and of Difference Score
plot_NumbLet_BIS <- ggplot(NumbLet_subset_BIS_long, aes(x=BIS)) + 
  geom_histogram(binwidth=0.5, colour="black", fill="white") + 
  facet_grid(Score ~ .) +
  labs(title = "Score Distribution Number-Letter Task by Condition",
       x = "Score",
       y = "Frequency")
```

# 13. Distribution plots for Stroop Task

## Distribution of RT for each condition and group
```{r Stroop: distribution of RT}

# Merge data from HC and patients, bring to long-format (to display in facet)
Stroop_combined <- rbind(data_Stroop_HC, data_Stroop_Pat_pre[,1:12])
Stroop_subset_RT <- Stroop_combined[c("Congruent_RT", "Incongruent_RT", "Gruppe")]
Stroop_subset_RT_long <- gather(Stroop_subset_RT, key = "Condition", value = "RT", Congruent_RT, Incongruent_RT)

# Distribution plot of RT for both conditions and groups
plot_Stroop_RT <- ggplot(Stroop_subset_RT_long, aes(x=RT, fill=as.factor(Gruppe))) + 
  geom_histogram(binwidth=25, colour="black", fill="white") + 
  facet_grid(Gruppe ~ Condition) +
  labs(title = "RT Distribution Stroop Task by Group and Condition",
       x = "Reaction Time",
       y = "Frequency")
```

## HC: Distribution of mean RT across all conditions
```{r Stroop: Distribution of RT for only HCs}
mean_RT_HC = mean(data_Stroop_HC$Overall_RT)

plot_Stroop_RT_HC <- ggplot(data_Stroop_HC, aes(x=Overall_RT)) +
  geom_histogram(binwidth=25, colour="black", fill="white") +
  geom_vline(xintercept = mean_RT_HC, 
             col = "red") +
  scale_x_continuous(breaks = seq(400, 1000, 200), lim = c(400, 1000)) +
  labs(title="Distribution of mean RT across conditions for HCs", 
       x="Reaction Time", 
       y="Frequency")
```

## Distribution of PC for each condition and group
```{r Stroop: distribution of PC}

# Subset and bring to long-format
Stroop_subset_PC <- Stroop_combined[c("Congruent_PC", "Incongruent_PC", "Gruppe")]
Stroop_subset_PC_long <- gather(Stroop_subset_PC, key = "Condition", value = "PC", Congruent_PC, Incongruent_PC)

# Distribution plot of PC for both conditions and groups
plot_Stroop_PC <- ggplot(Stroop_subset_PC_long, aes(x=PC, fill=as.factor(Gruppe))) + 
  geom_histogram(binwidth=0.04, colour="black", fill="white") + 
  facet_grid(Gruppe ~ Condition) +
  labs(title = "PC Distribution Stroop Task by Group and Condition",
       x = "Proportion Correct",
       y = "Frequency")
```

## HC: Distribution of mean PC across all conditions
```{r Stroop: Distribution of PC for only HCs}
mean_PC_HC = mean(data_Stroop_HC$Overall_PC)

plot_Stroop_PC_HC <- ggplot(data_Stroop_HC, aes(x=Overall_PC)) +
  geom_histogram(binwidth=0.04, colour="black", fill="white") +
  geom_vline(xintercept = mean_PC_HC, 
             col = "red") +
  scale_x_continuous(breaks = seq(0.2, 1.0, 0.2), lim = c(0.2, 1.1)) +
  labs(title="Distribution of mean PC across conditions for HCs", 
       x="Proportion Correct", 
       y="Frequency")
```

## Distribution of BIS for each condition 
```{r Stroop: Distribution of BIS & Diff_Score}

# Bring to long-format (to display in facet)
Stroop_subset_BIS <- data_Stroop_Pat_pre[c("BIS_Congruent", "BIS_Incongruent", "Diff_Score")]
Stroop_subset_BIS_long <- gather(Stroop_subset_BIS, key = "Score", value = "BIS", BIS_Congruent, BIS_Incongruent, Diff_Score)

# Distribution plot of BIS for both conditions and of Difference Score
plot_Stroop_BIS <- ggplot(Stroop_subset_BIS_long, aes(x=BIS)) + 
  geom_histogram(binwidth=0.5, colour="black", fill="white") + 
  facet_grid(Score ~ .) +
  labs(title = "Score Distribution Stroop Task by Condition",
       x = "Score",
       y = "Frequency")
```

# 14. Distribution plots for TwoBack Task

## Distribution of RT for each condition and group
```{r TwoBack: distribution of RT}

# Merge data from HC and patients, bring to long-format (to display in facet)
TwoBack_combined <- rbind(data_TwoBack_HC, data_TwoBack_Pat_pre[,1:12])
TwoBack_subset_RT <- TwoBack_combined[c("Overall_RT", "Target_RT", "Foil_RT", "Gruppe")]
TwoBack_subset_RT_long <- gather(TwoBack_subset_RT, key = "Condition", value = "RT", Overall_RT, Target_RT, Foil_RT)

# Distribution plot of RT for both conditions and groups
plot_TwoBack_RT <- ggplot(TwoBack_subset_RT_long, aes(x=RT, fill=as.factor(Gruppe))) + 
  geom_histogram(binwidth=40, colour="black", fill="white") + 
  facet_grid(Gruppe ~ Condition) +
  labs(title = "RT Distribution TwoBack Task by Group and Condition",
       x = "Reaction Time",
       y = "Frequency")
```

## HC: Distribution of mean RT across all conditions
```{r TwoBack: Distribution of RT for only HCs}
mean_RT_HC = mean(data_TwoBack_HC$Overall_RT)

plot_TwoBack_RT_HC <- ggplot(data_TwoBack_HC, aes(x=Overall_RT)) +
  geom_histogram(binwidth=40, colour="black", fill="white") +
  geom_vline(xintercept = mean_RT_HC, 
             col = "red") +
  scale_x_continuous(breaks = seq(250, 1000, 250), lim = c(250, 1125)) +
  labs(title="Distribution of mean RT across conditions for HCs", 
       x="Reaction Time", 
       y="Frequency")
```

## Distribution of PC for each condition and group
```{r TwoBack: distribution of PC}

# Subset and bring to long-format
TwoBack_subset_PC <- TwoBack_combined[c("Overall_PC", "Target_PC", "Foil_PC", "Gruppe")]
TwoBack_subset_PC_long <- gather(TwoBack_subset_PC, key = "Condition", value = "PC", Overall_PC, Target_PC, Foil_PC)

# Distribution plot of PC for both conditions and groups
plot_TwoBack_PC <- ggplot(TwoBack_subset_PC_long, aes(x=PC, fill=as.factor(Gruppe))) + 
  geom_histogram(binwidth=0.06, colour="black", fill="white") + 
  facet_grid(Gruppe ~ Condition) +
  labs(title = "PC Distribution TwoBack Task by Group and Condition",
       x = "Proportion Correct",
       y = "Frequency")
```

## HC: Distribution of mean PC across all conditions
```{r TwoBack: Distribution of PC for only HCs}
mean_PC_HC = mean(data_TwoBack_HC$Overall_PC)

plot_TwoBack_PC_HC <- ggplot(data_TwoBack_HC, aes(x=Overall_PC)) +
  geom_histogram(binwidth=0.06, colour="black", fill="white") +
  geom_vline(xintercept = mean_PC_HC, 
             col = "red") +
  scale_x_continuous(breaks = seq(0.0, 1.0, 0.25), lim = c(0.0, 1.1)) +
  labs(title="Distribution of mean PC across conditions for HCs", 
       x="Proportion Correct", 
       y="Frequency")
```

## Distribution of BIS for each condition 
```{r TwoBack: Distribution of BIS & Diff_Score}

# Bring to long-format (to display in facet)
TwoBack_subset_BIS <- data_TwoBack_Pat_pre[c("BIS_Total", "BIS_Target", "BIS_Foil")]
TwoBack_subset_BIS_long <- gather(TwoBack_subset_BIS, key = "Score", value = "BIS", BIS_Total, BIS_Target, BIS_Foil)

# Distribution plot of BIS for both conditions and of Difference Score
plot_TwoBack_BIS <- ggplot(TwoBack_subset_BIS_long, aes(x=BIS)) + 
  geom_histogram(binwidth=0.3, colour="black", fill="white") + 
  facet_grid(Score ~ .) +
  labs(title = "Score Distribution 2-Back Task by Condition",
       x = "Score",
       y = "Frequency")
```