---
title: "Group Comparison"
authors: "Charlotte Meinke, Rebecca Delfendahl, Till Julius Adam"
date: "2024-09-04"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---
<style type="text/css">
.main-container { /* Adjust main blocks */
  max-width: 100% !important;
  margin: auto;
}

body {
  font-family: "Georgia", serif !important; /* Set the font for the entire document */
}

.tocify { /* Adjust table of contents */
  max-width: 100% !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(pander)
library(effsize)
library(pwr)
library(gtsummary)
library(stats)
```


## <u>Importing Data</u>

### Task Performance Data
```{r}
data_HC <- read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/With_wrong_responses/Data_HC.csv")
data_Pat_pre <- read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/With_wrong_responses/Data_Patients_Pre.csv")
data_Pat_post <- read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/With_wrong_responses/Data_Patients_Post.csv")
```

### Subject Data
```{r}
data_all <- haven::read_dta('Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/CognitiveControl_Treatment/Data_Kevin_28.07.23.dta')
colnames(data_all)[which(colnames(data_all) == "id")] <- "Subject"
```

## <u>Comparison HC vs. Patients</u>

### Prepare dataset
```{r}
## 1. Merge patients pre and controls
Patients_vs_HC <- rbind(data_Pat_pre, data_HC)

## 2. Reduce dataset to BIS columns and SSRT
BIS_columns <- colnames(Patients_vs_HC)[grep("BIS", colnames(Patients_vs_HC))]
imp_columns <- c(BIS_columns, "SSRT")
Patients_vs_HC_imp <- Patients_vs_HC[,c("Subject","Gruppe",imp_columns)]

## 3. Add confounds (Alter, Geschlecht, Abschluss)
Patients_vs_HC_imp_conf <- merge(data_all[,c("Subject","Alter","Geschlecht","Abschluss")], Patients_vs_HC_imp, by = "Subject")
```
Data for `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1,])` patients and `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0,])` healthy controls were available.
Find assumption test for normality in script "Testing Normality"

### Calculate Welch t-test  
```{r}
t_test_mult_cols <- function(df_basis, cols) {
  df <- data.frame(t_statistic = numeric(length(cols)), p_value = numeric(length(cols)),
                   group_mean_Patients = numeric(length(cols)), sd_Patients = numeric(length(cols)),
                   group_mean_HC = numeric(length(cols)), sd_HC = numeric(length(cols)),
                   cohen_d = numeric(length(cols)))
  rownames(df) <- cols
  p_values_raw <- numeric(length(cols))  # Vector to store raw p-values
  
  for (i in seq_along(cols)) {
    col <- cols[i] #seq_along for sequencing column indices
    
    # Using the non-formula interface
    group0 <- na.omit(df_basis[df_basis[["Gruppe"]] == 0, col])
    group1 <- na.omit(df_basis[df_basis[["Gruppe"]] == 1, col])
    results <- t.test(group0, group1, paired = FALSE, var.equal = FALSE)
    #Alternativ using formula method:
      #results <- t.test(df_basis[[col]] ~ df_basis[["Gruppe"]], paired = FALSE, var.equal = FALSE)
    
    # Store raw p-values for BH correction
    p_values_raw[i] <- results$p.value
    
    # Calculate Cohen's d using the cohen.d function (effect size)
    cohen_d_result <- cohen.d(group0, group1, hedges.correction = FALSE)
    cohen_d <- cohen_d_result$estimate
    
    # Calculate power using pwr.t.test
    n0 <- length(group0)
    n1 <- length(group1)
    power_result <- pwr.t.test(d = cohen_d, n = min(n0, n1), sig.level = 0.05, type = "two.sample", alternative = "greater")
    power <- power_result$power
    
    # Store the results
    df[col, "t_statistic"] <- round(results$statistic, 2)
    df[col, "p_value"] <- round(results$p.value, 2)
    df[col, "group_mean_Patients"] <- round(mean(group1), 2)
    df[col, "sd_Patients"] <- round(sd(group0), 2)
    df[col, "group_mean_HC"] <- round(mean(group0), 2)
    df[col, "sd_HC"] <- round(sd(group1), 2)
    df[col, "cohen_d"] <- round(cohen_d, 2)
    df[col, "power"] <- round(power, 2)
  }
  
  # Adjust p-values using Benjamini-Hochberg method for multiple testing of related tasks
  p_values_adjusted <- p.adjust(p_values_raw, method = "BH")
  df$p_value_adjusted <- round(p_values_adjusted, 2)
  
  return(df)
}

t_test_table <- t_test_mult_cols(df_basis = Patients_vs_HC_imp, cols = imp_columns)
pander(t_test_table, style = "rmarkdown", fontsize = "tiny")
```
Patients performed worse than healthy controls, especially in the Stroop and the NumberLetter Task. However, please note that we have not tested for baseline differences (e.g., education, age...). This is done in the following step.

### Control Welch-test for covariates age, sex, and education
```{r}
#Descriptive statistics
table1 <- 
  Patients_vs_HC_imp_conf %>%
  tbl_summary(include = c(Alter, Geschlecht, Abschluss), by = Gruppe)
pander(as.data.frame(table1))
```

```{r}
#ANCOVA
ANCOVA_mult_cols <- function(df_basis, cols, covariates = NULL){
  # Create a overview of ANCOVA results for several columns in a dataframe
  df <- data.frame(Gruppe_p_value = numeric(length(cols)))
  rownames(df) <- cols
  for (col in cols){
    # Run ANCOVA
    formula <- paste(col, " ~ Gruppe", ifelse(length(covariates) > 0, paste(" + ", paste(covariates, collapse = " + ")), ""))
    anova_result <- aov(as.formula(formula), data = df_basis)
    coefficients <- coef(anova_result)
    summary_anova <- summary(anova_result)
    #print(summary_anova)
    
    # Prepare nice table to summarize the results
    df[col, "Gruppe_p_value"] <- round(summary_anova[[1]]$"Pr(>F)"[1],2)
    df[col, "Gruppe_corrected_difference"] <- round(coefficients[2],2)
    df[col, "Alter_p_value"] <- round(summary_anova[[1]]$"Pr(>F)"[2],2)
    df[col, "Geschlecht_p_value"] <- round(summary_anova[[1]]$"Pr(>F)"[3],2)
    df[col, "Abschluss_p_value"] <- round(summary_anova[[1]]$"Pr(>F)"[4],2)
  }
  return(df)
}

table_ANCOVA <- ANCOVA_mult_cols(df_basis = Patients_vs_HC_imp_conf, cols = imp_columns, covariates = c("Alter","Geschlecht","Abschluss"))
pander(table_ANCOVA, style = "rmarkdown", fontsize = "tiny")
```
3 observations with missings in the covariates are automatically deleted.
In the Numberletter task, the difference between patients and healthy controls remained when controlling for covariates. However, also education had a main effect on performance.
In the Stroop task, the difference between patients and healthy controls also remained when contrlloing for covariates. Here, also the gender and education had a main effect on performance.
There was no significant effect between patients and controls in the TwoBackTask - also not when controlling for covariates. The task types Total and Foil might mainly depend on the age.

## <u>Comparison pre vs.post</u>

### Pre-Post-Validation
```{r}
## Copy original dataframes
data_Pat_pre_comb <- data_Pat_pre
data_Pat_post_comb <- data_Pat_post

## Add timepoint-variable
data_Pat_pre_comb$time <- "pre"
data_Pat_post_comb$time <- "post"

# Combine data
data_Pat_pre_comb$Subject <- data_Pat_pre_comb$Subject + 600 # This was the ID given 
Patients_pre_vs_post <- rbind(data_Pat_post_comb, data_Pat_pre_comb)

# Create a summary table to inspect pre/post values for each patient
manual_validation <- Patients_pre_vs_post %>%
  group_by(Subject) %>%
  summarize(
    has_pre = any(time == "pre"),
    has_post = any(time == "post")
  )

# Display the summary table
print(manual_validation)

# Remove data if subjects had only pre-values
subject_counts <- table(Patients_pre_vs_post$Subject)
multiple_rows_subjects <- names(subject_counts)[subject_counts > 1]
Patients_pre_vs_post_clean <- Patients_pre_vs_post[Patients_pre_vs_post$Subject %in% multiple_rows_subjects, ]

# Reduce to BIS
#BIS_columns <- colnames(Patients_pre_vs_post_clean)[grep("BIS", colnames(Patients_pre_vs_post_clean))] # redundant
#imp_columns <- c(BIS_columns, "SSRT") # redundant
Patients_pre_vs_post_imp <- Patients_pre_vs_post_clean[,c("Subject",imp_columns,"time")]
```
Pre and post measures were available for `r nrow(Patients_pre_vs_post_imp) / 2` patients.

### Testing Assumption of dependent sample t-test: Positive correlation of measurement series 
# TODO: hardcode separation of Patients_pre_vs_post_imp dataframe into pre and post data (to avoid redundancy)
```{r}
#Helper function to avoid redundancy (Get pre and post data in one place for assumption and Welch test)
get_pre_post_data <- function(df_basis, col) {
  pre_data <- df_basis[df_basis[["time"]] == "pre", col]
  post_data <- df_basis[df_basis[["time"]] == "post", col]
  return(list(pre = pre_data, post = post_data))
}
```

```{r}
## Calculate Raw Pearson Correlation and Test Significance
calculate_correlations_and_significance <- function(df_basis, cols) {
  df <- data.frame(Correlation = numeric(length(cols)), t_statistic = numeric(length(cols)), 
                   p_value = numeric(length(cols)))
  rownames(df) <- cols
  p_values_raw <- numeric(length(cols))
  for (i in seq_along(cols)) {
    col <- cols[i]
    # Get pre and post data using the new function
    data <- get_pre_post_data(df_basis, col)
    # Use complete.cases to keep only rows where both pre AND post are not NA
    complete_cases <- complete.cases(data$pre, data$post)
    pre_data <- data$pre[complete_cases]
    post_data <- data$post[complete_cases]
    
    # Calculate Pearson correlation
    correlation <- cor(pre_data, post_data)
    
    # Calculate the t-statistic for the correlation
    n <- length(pre_data)
    t_statistic <- correlation * sqrt((n - 2) / (1 - correlation^2))
    
    # Calculate the p-value for a one-tailed test (H_0: r ≤ 0)
    p_value <- pt(t_statistic, df = n - 2, lower.tail = FALSE)
    p_values_raw[i] <- p_value
    
    # Store the results
    df[col, "Correlation"] <- round(correlation, 3)
    df[col, "t_statistic"] <- round(t_statistic, 3)
    df[col, "p_value"] <- round(p_value, 3)
  }
  
  p_values_adjusted <- p.adjust(p_values_raw, method = "BH")
  df$p_value_adjusted <- round(p_values_adjusted, 2)
  
  return(df)
}


# Apply the correlation and significance test to your data
correlation_significance_table <- calculate_correlations_and_significance(df_basis = Patients_pre_vs_post_imp, cols = imp_columns)
pander(correlation_significance_table, style = "rmarkdown", fontsize = "tiny")
```
For all tasks, the correlations of the measurement series are significantly positive. Thus, the assumption is met for all tasks. 

### Calculate Welch t-test
```{r}
paired_t_test_mult_cols <- function(df_basis, cols) {
  df <- data.frame(t_statistic = numeric(length(cols)), p_value = numeric(length(cols)),
                   group_mean_pre = numeric(length(cols)), group_mean_post = numeric(length(cols)),
                   cohen_d = numeric(length(cols)))
  rownames(df) <- cols
  p_values_raw <- numeric(length(cols)) # Vector to store raw p-values
  for (i in seq_along(cols)) {
    col <- cols[i]
    # Get pre and post data using get_pre_post_data helper function
    data <- get_pre_post_data(df_basis, col)
    # Use complete.cases to keep only rows where both pre AND post are not NA
    complete_cases <- complete.cases(data$pre, data$post)
    pre_data <- data$pre[complete_cases]
    post_data <- data$post[complete_cases]
    
    # non-formula interface as before
    results <- t.test(pre_data, post_data, paired = TRUE, var.equal = FALSE)
    #Alternativ using formula method:
      #results <- t.test(df_basis[[col]] ~ df_basis[["time"]], paired = TRUE, var.equal = FALSE)
    
    # Store raw p-values for BH correction
    p_values_raw[i] <- results$p.value
    
    # Calculate Cohen's d using the cohen.d function for paired samples (effect size)
    cohen_d_result <- cohen.d(post_data, pre_data, paired = TRUE)
    cohen_d <- cohen_d_result$estimate
    
    # Calculate power using pwr.t.test
    n <- length(pre_data)
    power_result <- pwr.t.test(d = cohen_d, n = n, sig.level = 0.05, type = "paired", alternative = "greater")
    power <- power_result$power
    
    # Store the results
    df[col, "t_statistic"] <- round(results$statistic, 2)
    df[col, "p_value"] <- round(results$p.value, 2)
    df[col, "group_mean_pre"] <- round(mean(pre_data), 2)
    df[col, "sd_pre"] <- round(sd(pre_data), 2)
    df[col, "group_mean_post"] <- round(mean(post_data), 2)
    df[col, "sd_post"] <- round(sd(post_data), 2)
    df[col, "cohen_d"] <- round(cohen_d, 2)
    df[col, "power"] <- round(power, 2)
  }
  
  # Adjust p-values using Benjamini-Hochberg method for multiple testing of related tasks
  p_values_adjusted <- p.adjust(p_values_raw, method = "BH")
  df$p_value_adjusted <- round(p_values_adjusted, 2)
  
  return(df)
}

t_test_table_pre_post <- paired_t_test_mult_cols(df_basis = Patients_pre_vs_post_imp, cols = imp_columns)
pander(t_test_table_pre_post)
```
Patients performed better in all measurements of executive functioning after the exposure session. This might be due to a practice effect.

## <u>Plotting</u>

### Comparison HC vs. patients (pre)
# TODO: include SSRT or do separate plot?
```{r}
# Reshape data from wide to long format
Patients_vs_HC_long <- Patients_vs_HC_imp %>%
  pivot_longer(cols = all_of(BIS_columns), # imp_columns to include SSRT
               names_to = "Condition",
               values_to = "BIS_Score")

# Add new task variable to keep the headers intact for faceting
Patients_vs_HC_long <- Patients_vs_HC_long %>%
  mutate(Task = case_when(
    grepl("NumberLetter", Condition) ~ "NumberLetter",
    grepl("Stroop", Condition) ~ "Stroop",
    grepl("TwoBack", Condition) ~ "TwoBack"
  ))

# Remove prefixes from Condition labels, keeping Task intact
Patients_vs_HC_long <- Patients_vs_HC_long %>%
  mutate(Condition = gsub("NumberLetter_BIS_|Stroop_BIS_|TwoBack_BIS_", "", Condition))

# Convert "Gruppe" to factor # TODO: change order of the conditions for consistency
Patients_vs_HC_long$Gruppe <- factor(Patients_vs_HC_long$Gruppe, levels = c(0, 1), labels = c("Healthy Controls", "Patients"))

# Create boxplots
boxplots_Patients_HC <- ggplot(Patients_vs_HC_long, aes(x = Condition, y = BIS_Score, color = Gruppe)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(position = position_jitterdodge(),
              size = 0.7,
              shape = 21,
              fill = NA) +
  facet_wrap(~ Task, scales = "free_x") +
  #facet_grid(rows = vars(Task), scales = "free_x") +
  labs(title = "Boxplots of Task Condition by Group",
       x = "Condition",
       y = "BIS-Score",
       color = "Group")

# Create violin plots with mean
violin_plots_patients_hc <- ggplot(Patients_vs_HC_long, aes(x = Condition, y = BIS_Score, fill = Gruppe)) +
  geom_violin(position = position_dodge(width = 0.8)) +
  stat_summary(fun.data = mean_sdl, geom = "pointrange", # mult = 2 (SD) by default
               position = position_dodge(width = 0.8)) + # width = 0.65, fatten = 2
  facet_wrap(~ Task, scales = "free_x") +
  labs(title = "Violin Plots of Task Condition by Group",
       x = "Task Condition",
       y = "BIS-Score",
       fill = "Group")
```

### Boxplot: Patients vs. Healthy Controls
```{r}
print(boxplots_Patients_HC)
```

### Violin Plot: Patients vs. Healthy Controls
```{r}
print(violin_plots_patients_hc)
```

### Comparison patients pre vs. post
# TODO: include SSRT or do separate plot?
```{r}
# Reshape data from wide to long format
Patients_pre_vs_post_long <- Patients_pre_vs_post_imp %>%
  pivot_longer(cols = all_of(BIS_columns), # imp_columns to include SSRT
               names_to = "Condition",
               values_to = "BIS_Score")

# Add new task variable
Patients_pre_vs_post_long <- Patients_pre_vs_post_long %>%
  mutate(Task = case_when(
    grepl("NumberLetter", Condition) ~ "NumberLetter",
    grepl("Stroop", Condition) ~ "Stroop",
    grepl("TwoBack", Condition) ~ "TwoBack"
  ))

# Remove prefixes from Condition labels, keeping Task intact
Patients_pre_vs_post_long <- Patients_pre_vs_post_long %>%
  mutate(Condition = gsub("NumberLetter_BIS_|Stroop_BIS_|TwoBack_BIS_", "", Condition))

# TODO: change order of the conditions for consistency
# Convert "time" to factor, specify order
Patients_pre_vs_post_long$time <- factor(Patients_pre_vs_post_long$time, levels = c("pre", "post"))

# Create boxplot with paired observations
boxplot_Patients_pre_post <- ggplot(Patients_pre_vs_post_long, aes(x = Condition, y = BIS_Score, color = time)) + # group = Subject
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(position = position_jitterdodge(),
              size = 0.7,
              shape = 21,
              fill = NA) +
  #geom_line(aes(group = Subject)) +
  facet_wrap(~ Task, scales = "free_x") +
  labs(title = "Boxplots of Task Condition by Time of Measurement",
       x = "Condition",
       y = "BIS-Score",
       color = "Time")
```

### Boxplot for Patients pre vs. post intervention
```{r}
print(boxplot_Patients_pre_post)
```