---
title: "Comparison_EF"
author: "Charlotte Meinke"
date: "2024-01-24"
output: 
  html_document:
      df_print: paged
      toc: yes
      toc_float: true
      toc_depth: 3
      number_sections: yes
      code_folding: hide
      fig_caption: yes
---

# Set-up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(pander)
```

# Import data 
## Import task performance data
```{r}
NumberLetter_HC <- read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/NumberLetter_HC.csv")
NumberLetter_Patients_Post <- read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/NumberLetter_Patients_Post.csv")
NumberLetter_Patients_Pre <-  read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/NumberLetter_Patients_Pre.csv")

Stroop_HC <- read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/Stroop_HC.csv")
Stroop_Patients_Post <-  read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/Stroop_Patients_Post.csv")
Stroop_Patients_Pre <-  read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/Stroop_Patients_Pre.csv")

TwoBack_HC <- read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/TwoBack_HC.csv")
TwoBack_Patients_Post <-  read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/TwoBack_Patients_Post.csv")
TwoBack_Patients_Pre <-  read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/TwoBack_Patients_Pre.csv")
```

## Import 
```{r}
data_all <- haven::read_dta('Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/CognitiveControl_Treatment/Data_Kevin_28.07.23.dta')
colnames(data_all)[which(colnames(data_all) == "id")] <- "Subject"
```

# Comparison HC vs. Patients

## Prepare dataset
```{r}
# Merge patients and controls
## 1. Merge patients Pre (different tasks)
Patients_pre_2 <- merge(NumberLetter_Patients_Pre, Stroop_Patients_Pre, by = "Subject")
Patients_pre_3 <- merge(Patients_pre_2, TwoBack_Patients_Pre, by = "Subject")

## 2. Merge controls (different tasks)
Controls_2 <- merge(NumberLetter_HC, Stroop_HC, by = "Subject")
Controls_3 <- merge(Controls_2, TwoBack_HC, by = "Subject")

## 3. Merge Controls and patients
Patients_vs_HC <- rbind(Patients_pre_3, Controls_3)

## 4. Reduce dataset to BIS columns
BIS_columns <- colnames(Patients_vs_HC)[startsWith(colnames(Patients_vs_HC), "BIS")]
Patients_vs_HC_imp <- Patients_vs_HC[,c("Subject",BIS_columns,"Gruppe")]

## 5. Add better names 
Patients_vs_HC_imp <- Patients_vs_HC_imp %>%
  rename(NumberLetter_BIS_Repeat = BIS_Repeat,
         NumberLetter_BIS_Switch = BIS_Switch,
         Stroop_BIS_Congruent = BIS_Congruent,
         Stroop_BIS_Incongruent = BIS_Incongruent,
         TwoBack_BIS_Total = BIS_Total,
         TwoBack_BIS_Target = BIS_Target,
         TwoBack_BIS_Foil = BIS_Foil)

BIS_columns_renamed <- grep("BIS", colnames(Patients_vs_HC_imp), value = TRUE)

## 6. Add confounds (Alter, Geschlecht, Abschluss)
Patients_vs_HC_imp_conf <- merge(data_all[,c("Subject","Alter","Geschlecht","Abschluss")], Patients_vs_HC_imp, by = "Subject")
```
Data for `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1,])` patients and `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0,])` healthy controls were available.


## Calculate Welch t-test  
```{r}
t_test_mult_cols <- function(df_basis, cols){
  # Create an overview of t-test results for several columns in a dataframe
  df <- data.frame(t_statistic = numeric(length(cols)), p_value = numeric(length(cols)))
rownames(df) <- cols
  for (col in cols){
    results <- t.test(df_basis[[col]] ~ df_basis[["Gruppe"]], paired = FALSE, var.equal = FALSE)
    df[col, "t_statistic"] <- round(results$statistic,2)
    df[col, "p_value"] <- round(results$p.value,2)
    df[col, "group_mean_Patients"] <- round(mean(df_basis[df_basis[["Gruppe"]] == 0, col]),2)
    df[col, "sd_Patients"] <- round(sd(df_basis[df_basis[["Gruppe"]] == 0, col]),2)
    df[col, "group_mean_HC"] <- round(mean(df_basis[df_basis[["Gruppe"]] == 1, col]),2)
    df[col, "sd_HC"] <- round(sd(df_basis[df_basis[["Gruppe"]] == 1, col]),2)
  }
  return(df)
}

t_test_table <- t_test_mult_cols(df_basis = Patients_vs_HC_imp, cols = BIS_columns_renamed)
pander(t_test_table, style = "rmarkdown", fontsize = "tiny")
```
Patients performed better than healthy controls, especially in the Stroop and the NumberLetterTask. However, please note that we have not tested for baseline differences (e.g., education, age...). This is done in the following step

## Control Welch-test for covariates age, sex, and education
```{r}
library(gtsummary)
table1 <- 
  Patients_vs_HC_imp_conf %>%
  tbl_summary(include = c(Alter, Geschlecht, Abschluss), by = Gruppe)

table1
```


```{r}
ANCOVA_mult_cols <- function(df_basis, cols, covariates = NULL){
  # Create a overview of ANCOVA results for several columns in a dataframe
  df <- data.frame(Gruppe_p_value = numeric(length(cols)))
  rownames(df) <- cols
  for (col in cols){
    # Run ANCOVA
    formula <- paste(col, " ~ Gruppe", ifelse(length(covariates) > 0, paste(" + ", paste(covariates, collapse = " + ")), ""))
    anova_result <- aov(as.formula(formula), data = df_basis)
    coefficients <- coef(anova_result)
    summary_anova <- summary(anova_result)
    print(summary_anova)
    
    # Prepare nice table to summarize the results
    df[col, "Gruppe_p_value"] <- round(summary_anova[[1]]$"Pr(>F)"[1],2)
    df[col, "Gruppe_corrected_difference"] <- round(coefficients[2],2)
    df[col, "Alter_p_value"] <- round(summary_anova[[1]]$"Pr(>F)"[2],2)
    df[col, "Geschlecht_p_value"] <- round(summary_anova[[1]]$"Pr(>F)"[3],2)
    df[col, "Abschluss_p_value"] <- round(summary_anova[[1]]$"Pr(>F)"[4],2)
  }
  return(df)
}

table_ANCOVA <- ANCOVA_mult_cols(df_basis = Patients_vs_HC_imp_conf, cols = BIS_columns_renamed, covariates = c("Alter","Geschlecht","Abschluss"))
pander(table_ANCOVA, style = "rmarkdown", fontsize = "tiny")
```
3 observations with missings in the covariates are automatically deleted.
In the Numberletter task, the difference between patients and healthy controls remained when controlling for covariates. However, also education had a main effect on performance.
In the Stroop task, the difference between patients and healthy controls also remained when contrlloing for covariates. Here, also the gender and education had a main effect on performance.
There was no significant effect between patients and controls in the TwoBackTask - also not when controlling for covariates. The task types Total and Foil might mainly depend on the age.

# Comparison pre vs.post
```{r}
## Merge patients pre
Patients_pre_2 <- merge(NumberLetter_Patients_Pre, Stroop_Patients_Pre, by = "Subject")
Patients_pre_3 <- merge(Patients_pre_2, TwoBack_Patients_Pre, by = "Subject")
Patients_pre_3$time <- "pre"

## Merge patients Post
Patients_post_2 <- merge(NumberLetter_Patients_Post, Stroop_Patients_Post, by = "Subject")
Patients_post_3 <- merge(Patients_post_2, TwoBack_Patients_Post, by = "Subject")
Patients_post_3$time <- "post"

# Combine data
Patients_pre_3$Subject <- Patients_pre_3$Subject + 600 # This was the ID given 
Patients_pre_vs_post <- rbind(Patients_post_3, Patients_pre_3)
# Remove data if subjects had only pre-values
subject_counts <- table(Patients_pre_vs_post$Subject)
multiple_rows_subjects <- names(subject_counts)[subject_counts > 1]
Patients_pre_vs_post_clean <- Patients_pre_vs_post[Patients_pre_vs_post$Subject %in% multiple_rows_subjects, ]

# Reduce to BIS
BIS_columns <- colnames(Patients_pre_vs_post_clean)[startsWith(colnames(Patients_pre_vs_post_clean), "BIS")]
Patients_pre_vs_post_imp <- Patients_pre_vs_post_clean[,c("Subject",BIS_columns,"time")]

## Add better names 
Patients_pre_vs_post_imp <- Patients_pre_vs_post_imp %>%
  rename(NumberLetter_BIS_Repeat = BIS_Repeat,
         NumberLetter_BIS_Switch = BIS_Switch,
         Stroop_BIS_Congruent = BIS_Congruent,
         Stroop_BIS_Incongruent = BIS_Incongruent,
         TwoBack_BIS_Total = BIS_Total,
         TwoBack_BIS_Target = BIS_Target,
         TwoBack_BIS_Foil = BIS_Foil)
BIS_columns_renamed <- grep("BIS", colnames(Patients_pre_vs_post_imp), value = TRUE)
```
Pre and post measures were available for `r nrow(Patients_pre_vs_post_imp) / 2` patients.

## Calculate Welch t-test
```{r}
paired_t_test_mult_cols <- function(df_basis, cols){
  # Create a overview of t-test results for several columns in a dataframe
  df <- data.frame(t_statistic= numeric(length(cols)), p_value = numeric(length(cols)),
                   group_mean_pre = numeric(length(cols)), group_mean_post = numeric(length(cols)))
  rownames(df) <- cols
  for (col in cols){
    results <- t.test(df_basis[[col]] ~ df_basis[["time"]], paired = TRUE, var.equal = FALSE)
    df[col, "t_statistic"] <- round(results$statistic, 2)
    df[col, "p_value"] <- results$p.value
    # Calculate group means
    df[col, "group_mean_pre"] <- round(mean(df_basis[df_basis[["time"]] == "pre", col]),2)
    df[col, "sd_pre"] <- round(sd(df_basis[df_basis[["time"]] == "pre", col]),2)
    df[col, "group_mean_post"] <- round(mean(df_basis[df_basis[["time"]] == "post", col]),2)
    df[col, "sd_post"] <- round(sd(df_basis[df_basis[["time"]] == "post", col]),2)
  }
  return(df)
}

t_test_table_pre_post <- paired_t_test_mult_cols(df_basis = Patients_pre_vs_post_imp, cols = BIS_columns_renamed)
pander(t_test_table_pre_post)
t_test_table_pre_post
```
Patients performed better in all measurements of executive functioning after the exposure session. This might be due to a practice effect.

# Plotting

## Comparison HC vs. patients (pre)
```{r}
# Reshape data from wide to long format
Patients_vs_HC_long <- Patients_vs_HC_imp %>%
  pivot_longer(cols = all_of(BIS_columns_renamed),
               names_to = "Condition",
               values_to = "BIS_Score")

# Add new task variable
Patients_vs_HC_long <- Patients_vs_HC_long %>%
  mutate(Task = case_when(
    grepl("NumberLetter", Condition) ~ "NumberLetter",
    grepl("Stroop", Condition) ~ "Stroop",
    grepl("TwoBack", Condition) ~ "TwoBack"
  ))

# Convert "Gruppe" to factor
Patients_vs_HC_long$Gruppe <- factor(Patients_vs_HC_long$Gruppe)

# Create boxplots # TODO: try violin-plot instead
boxplots_Patients_HC <- ggplot(Patients_vs_HC_long, aes(x = Condition, y = BIS_Score, color = Gruppe)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(position = position_jitterdodge(),
              size = 0.7,
              shape = 21,
              fill = NA) +
  facet_wrap(~ Task, scales = "free_x") +
  #facet_grid(rows = vars(Task), scales = "free_x") +
  labs(title = "Boxplots of Task Condition by Group",
       x = "Condition",
       y = "BIS-Score",
       color = "Group")

# Create violin plots with mean
violin_plots_patients_hc <- ggplot(Patients_vs_HC_long, aes(x = Condition, y = BIS_Score, fill = Gruppe)) +
  geom_violin(position = position_dodge(width = 0.8)) +
  stat_summary(fun.data = mean_sdl, geom = "pointrange", # mult = 2 (SD) by default
               position = position_dodge(width = 0.8)) + # width = 0.65, fatten = 2
  facet_wrap(~ Task, scales = "free_x") +
  labs(title = "Violin Plots of Task Condition by Group",
       x = "Task Condition",
       y = "BIS-Score",
       fill = "Group")
```

## Comparison patients pre vs. post
```{r}
# Reshape data from wide to long format
Patients_pre_vs_post_long <- Patients_pre_vs_post_imp %>%
  pivot_longer(cols = all_of(BIS_columns_renamed),
               names_to = "Condition",
               values_to = "BIS_Score")

# Add new task variable
Patients_pre_vs_post_long <- Patients_pre_vs_post_long %>%
  mutate(Task = case_when(
    grepl("NumberLetter", Condition) ~ "NumberLetter",
    grepl("Stroop", Condition) ~ "Stroop",
    grepl("TwoBack", Condition) ~ "TwoBack"
  ))

# Convert "time" to factor, specify order
Patients_pre_vs_post_long$time <- factor(Patients_pre_vs_post_long$time, levels = c("pre", "post"))

# Create boxplot with paired observations
boxplot_Patients_pre_post <- ggplot(Patients_pre_vs_post_long, aes(x = Condition, y = BIS_Score, color = time)) + # group = Subject
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(position = position_jitterdodge(),
              size = 0.7,
              shape = 21,
              fill = NA) +
  #geom_line(aes(group = Subject)) +
  facet_wrap(~ Task, scales = "free_x") +
  labs(title = "Boxplots of Task Condition by Time of Measurement",
       x = "Condition",
       y = "BIS-Score",
       color = "Time")
```

