---
title: "Testing Normality"
author: "Till Julius Adam"
date: "2024-09-04"
output: html_document
---

# Set-up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(ggplot2)
library(dplyr)
library(pander)
```

# Import data 
## Import task performance data
```{r}
NumberLetter_HC <- read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/NumberLetter_HC.csv")
NumberLetter_Patients_Post <- read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/NumberLetter_Patients_Post.csv")
NumberLetter_Patients_Pre <-  read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/NumberLetter_Patients_Pre.csv")

Stroop_HC <- read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/Stroop_HC.csv")
Stroop_Patients_Post <-  read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/Stroop_Patients_Post.csv")
Stroop_Patients_Pre <-  read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/Stroop_Patients_Pre.csv")

TwoBack_HC <- read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/TwoBack_HC.csv")
TwoBack_Patients_Post <-  read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/TwoBack_Patients_Post.csv")
TwoBack_Patients_Pre <-  read.csv("Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Daten_Gruppenvergleich/TwoBack_Patients_Pre.csv")
```

## Import 
```{r}
data_all <- haven::read_dta('Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/CognitiveControl_Treatment/Data_Kevin_28.07.23.dta')
colnames(data_all)[which(colnames(data_all) == "id")] <- "Subject"
```

# Comparison HC vs. Patients

## Prepare dataset
```{r}
# Merge patients and controls
## 1. Merge patients Pre (different tasks)
Patients_pre_2 <- merge(NumberLetter_Patients_Pre, Stroop_Patients_Pre, by = "Subject")
Patients_pre_3 <- merge(Patients_pre_2, TwoBack_Patients_Pre, by = "Subject")

## 2. Merge controls (different tasks)
Controls_2 <- merge(NumberLetter_HC, Stroop_HC, by = "Subject")
Controls_3 <- merge(Controls_2, TwoBack_HC, by = "Subject")

## 3. Merge Controls and patients
Patients_vs_HC <- rbind(Patients_pre_3, Controls_3)

## 4. Reduce dataset to BIS columns
BIS_columns <- colnames(Patients_vs_HC)[startsWith(colnames(Patients_vs_HC), "BIS")]
Patients_vs_HC_imp <- Patients_vs_HC[,c("Subject",BIS_columns,"Gruppe")]

## 5. Add better names 
Patients_vs_HC_imp <- Patients_vs_HC_imp %>%
  rename(NumberLetter_BIS_Repeat = BIS_Repeat,
         NumberLetter_BIS_Switch = BIS_Switch,
         Stroop_BIS_Congruent = BIS_Congruent,
         Stroop_BIS_Incongruent = BIS_Incongruent,
         TwoBack_BIS_Total = BIS_Total,
         TwoBack_BIS_Target = BIS_Target,
         TwoBack_BIS_Foil = BIS_Foil)

BIS_columns_renamed <- grep("BIS", colnames(Patients_vs_HC_imp), value = TRUE)

## 6. Add confounds (Alter, Geschlecht, Abschluss)
Patients_vs_HC_imp_conf <- merge(data_all[,c("Subject","Alter","Geschlecht","Abschluss")], Patients_vs_HC_imp, by = "Subject")
```
Data for `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1,])` patients and `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0,])` healthy controls were available.

## Testing assumption of independent sample t-test: normal distributions of the characteristic in both populations
#Implemented using a Kolmogorov-Smirnov test
```{r}
## Kolmogorov-Smirnov Test for Normality
ks_test_mult_cols <- function(df_basis, cols) {
  df <- data.frame(KS_statistic_group0 = numeric(length(cols)), p_value_group0 = numeric(length(cols)),
                   KS_statistic_group1 = numeric(length(cols)), p_value_group1 = numeric(length(cols)))
  rownames(df) <- cols
  for (col in cols) {
    group0 <- df_basis[df_basis[["Gruppe"]] == 0, col]
    group1 <- df_basis[df_basis[["Gruppe"]] == 1, col]
    
    # Perform Kolmogorov-Smirnov test comparing group data to a normal distribution
    ks_test_group0 <- ks.test(group0, "pnorm", mean(group0), sd(group0))
    ks_test_group1 <- ks.test(group1, "pnorm", mean(group1), sd(group1))
    
    # Store the results
    df[col, "KS_statistic_group0"] <- round(ks_test_group0$statistic, 3)
    df[col, "p_value_group0"] <- round(ks_test_group0$p.value, 3)
    df[col, "KS_statistic_group1"] <- round(ks_test_group1$statistic, 3)
    df[col, "p_value_group1"] <- round(ks_test_group1$p.value, 3)
  }
  return(df)
}

# Apply the K-S test to your data
ks_test_table <- ks_test_mult_cols(df_basis = Patients_vs_HC_imp, cols = BIS_columns_renamed)
pander(ks_test_table, style = "rmarkdown", fontsize = "tiny")

```
The Kolmogoroff-Smirnov tests show significant deviations from normal distribution for the Number-Letter and Stroop tasks in the healthy controls (alpha=0.05). However, with Bonferroni correction (alpha=0.003), only the Stroop test would violate the assumption.

#Visual Inspection
#Value Range
#In case one wants to manually set the value range oriented on min and max values
```{r}
get_min_max_values <- function(df_basis, cols) {
  do.call(rbind, lapply(cols, function(col) {
    do.call(rbind, lapply(unique(df_basis$Gruppe), function(group) {
      group_data <- df_basis[df_basis$Gruppe == group, col]
      data.frame(
        Group = ifelse(group == 0, "HC", "Patients"),
        Condition = col,
        Min_Value = min(group_data, na.rm = TRUE),
        Max_Value = max(group_data, na.rm = TRUE)
      )
    }))
  }))
}

min_max_values_table <- get_min_max_values(df_basis = Patients_vs_HC_imp, cols = BIS_columns_renamed)
pander(min_max_values_table, style = "rmarkdown", caption = "min max")
```
# In this case: -11.88 - 3.484


Manual Range
```{r}
create_density_plot <- function(data, col, group) {
  group_data <- data[data$Gruppe == group, col]
  
  ggplot(data.frame(x = group_data), aes(x = x)) +
    geom_density(fill = ifelse(group == 0, "blue", "red"), alpha = 0.5) +
    labs(title = paste(col, "- Group", ifelse(group == 0, "HC", "Patients")),
         x = col,
         y = "Density") +
    xlim(-10, 5) #Manual range based on biggest and smallest values
}

generate_density_plots <- function(df_basis, cols) {
  plots <- list()
  for (col in cols) {
    for (group in unique(df_basis$Gruppe)) {
      plots[[paste(col, group)]] <- create_density_plot(df_basis, col, group)
    }
  }
  return(plots)
}

density_plots <- generate_density_plots(df_basis = Patients_vs_HC_imp, cols = BIS_columns_renamed)
for (plot_name in names(density_plots)) {
  print(density_plots[[plot_name]])
}

save_directory <- "Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Density Plots_HCvsPat/Manual"
for (plot_name in names(density_plots)) {
  file_name <- paste0(save_directory, "/", plot_name, ".png")
  ggsave(file_name, plot = density_plots[[plot_name]])
}
```

Automatic whole range (automatically takes min and max values and adjusts value range automatically)
```{r}
global_x_range <- range(Patients_vs_HC_imp[, BIS_columns_renamed], na.rm = TRUE) #x-range as smallest and largest value for each group, automatic

create_density_plot <- function(data, col, group, x_range) {
  group_data <- data[data$Gruppe == group, col]
  
  ggplot(data.frame(x = group_data), aes(x = x)) +
    geom_density(fill = ifelse(group == 0, "blue", "red"), alpha = 0.5) +
    labs(title = paste(col, "- Group", ifelse(group == 0, "HC", "Patients")),
         x = col,
         y = "Density") +
    xlim(x_range[1], x_range[2])
}

generate_density_plots <- function(df_basis, cols, x_range) {
  plots <- list()
  for (col in cols) {
    for (group in unique(df_basis$Gruppe)) {
      plots[[paste(col, group)]] <- create_density_plot(df_basis, col, group, x_range)
    }
  }
  return(plots)
}

density_plots <- generate_density_plots(df_basis = Patients_vs_HC_imp, cols = BIS_columns_renamed, x_range = global_x_range)
for (plot_name in names(density_plots)) {
  print(density_plots[[plot_name]])
}

save_directory <- "Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Density Plots_HCvsPat/Automatic whole range"
for (plot_name in names(density_plots)) {
  file_name <- paste0(save_directory, "/", plot_name, ".png")
  ggsave(file_name, plot = density_plots[[plot_name]])
}
```


Automatic group range (Individual x-axis for each group according to min and max values of that task)
```{r}
create_density_plot <- function(data, col, group) {
  group_data <- data[data$Gruppe == group, col]
  x_range <- range(group_data, na.rm = TRUE) #By each group min max
  
  ggplot(data.frame(x = group_data), aes(x = x)) +
    geom_density(fill = ifelse(group == 0, "blue", "red"), alpha = 0.5) +
    labs(title = paste(col, "- Group", ifelse(group == 0, "HC", "Patients")),
         x = col,
         y = "Density") +
    xlim(x_range[1], x_range[2])
}

generate_density_plots <- function(df_basis, cols) {
  plots <- list()
  for (col in cols) {
    for (group in unique(df_basis$Gruppe)) {
      plots[[paste(col, group)]] <- create_density_plot(df_basis, col, group)
    }
  }
  return(plots)
}

density_plots <- generate_density_plots(df_basis = Patients_vs_HC_imp, cols = BIS_columns_renamed)
for (plot_name in names(density_plots)) {
  print(density_plots[[plot_name]])
}

save_directory <- "Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Density Plots_HCvsPat/Automatic group range"
for (plot_name in names(density_plots)) {
  file_name <- paste0(save_directory, "/", plot_name, ".png")
  ggsave(file_name, plot = density_plots[[plot_name]])
}
```



# Comparison pre vs.post
# data preparation pre post
```{r}
## Merge patients pre
Patients_pre_2 <- merge(NumberLetter_Patients_Pre, Stroop_Patients_Pre, by = "Subject")
Patients_pre_3 <- merge(Patients_pre_2, TwoBack_Patients_Pre, by = "Subject")
Patients_pre_3$time <- "pre"

## Merge patients Post
Patients_post_2 <- merge(NumberLetter_Patients_Post, Stroop_Patients_Post, by = "Subject")
Patients_post_3 <- merge(Patients_post_2, TwoBack_Patients_Post, by = "Subject")
Patients_post_3$time <- "post"

# Combine data
Patients_pre_3$Subject <- Patients_pre_3$Subject + 600 # This was the ID given 
Patients_pre_vs_post <- rbind(Patients_post_3, Patients_pre_3)
# Remove data if subjects had only pre-values
subject_counts <- table(Patients_pre_vs_post$Subject)
multiple_rows_subjects <- names(subject_counts)[subject_counts > 1]
Patients_pre_vs_post_clean <- Patients_pre_vs_post[Patients_pre_vs_post$Subject %in% multiple_rows_subjects, ]

# Reduce to BIS
BIS_columns <- colnames(Patients_pre_vs_post_clean)[startsWith(colnames(Patients_pre_vs_post_clean), "BIS")]
Patients_pre_vs_post_imp <- Patients_pre_vs_post_clean[,c("Subject",BIS_columns,"time")]

## Add better names 
Patients_pre_vs_post_imp <- Patients_pre_vs_post_imp %>%
  rename(NumberLetter_BIS_Repeat = BIS_Repeat,
         NumberLetter_BIS_Switch = BIS_Switch,
         Stroop_BIS_Congruent = BIS_Congruent,
         Stroop_BIS_Incongruent = BIS_Incongruent,
         TwoBack_BIS_Total = BIS_Total,
         TwoBack_BIS_Target = BIS_Target,
         TwoBack_BIS_Foil = BIS_Foil)
BIS_columns_renamed <- grep("BIS", colnames(Patients_pre_vs_post_imp), value = TRUE)
```
Pre and post measures were available for `r nrow(Patients_pre_vs_post_imp) / 2` patients.


## Testing assumption of dependent sample t-test: normal distributions of difference values post-pre
#Implemented using a Kolmogorov-Smirnov test on difference values (d_i=Post_i-Pre_i)
```{r}
## Kolmogorov-Smirnov Test for Normality of Difference Values
ks_test_diff_values <- function(df_basis, cols) {
  df <- data.frame(KS_statistic = numeric(length(cols)), p_value = numeric(length(cols)))
  rownames(df) <- cols
  for (col in cols) {
    # Calculate difference values (Post - Pre)
    diff_values <- df_basis[df_basis[["time"]] == "post", col] - df_basis[df_basis[["time"]] == "pre", col]
    ks_test_diff <- ks.test(diff_values, "pnorm", mean(diff_values), sd(diff_values))
    
    df[col, "KS_statistic"] <- round(ks_test_diff$statistic, 3)
    df[col, "p_value"] <- round(ks_test_diff$p.value, 3)
  }
  return(df)
}

ks_test_diff_table <- ks_test_diff_values(df_basis = Patients_pre_vs_post_imp, cols = BIS_columns_renamed)
pander(ks_test_diff_table, style = "rmarkdown", fontsize = "tiny")
```
The assumption of normal distribution is violated for all but the TwoBack task for alpha=0.05. 
However, if we apply Bonferroni correction (alpha=0.007) only the Number-Letter Repeat and Stroop tests would violate the assumption.

#Visual inspection
#Automatic group range
```{r}
create_density_plot_diff <- function(data, col) {
  # Calculate difference values (Post - Pre)
  diff_values <- data[data[["time"]] == "post", col] - data[data[["time"]] == "pre", col]
  x_range <- range(diff_values, na.rm = TRUE) #Group range
  
  ggplot(data.frame(x = diff_values), aes(x = x)) +
    geom_density(fill = "red", alpha = 0.5) +
    labs(title = paste("Difference:", col),
         x = paste("Difference in", col),
         y = "Density") +
    xlim(x_range[1], x_range[2])
}

density_plots_diff <- lapply(BIS_columns_renamed, function(col) {
  create_density_plot_diff(Patients_pre_vs_post_imp, col)
})

names(density_plots_diff) <- BIS_columns_renamed
for (plot in density_plots_diff) {
  print(plot)
}

save_directory <- "Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Density Plot_PrePost/Automatic group range"
for (plot_name in names(density_plots_diff)) {
  file_name <- paste0(save_directory, "/", plot_name, ".png")
  ggsave(file_name, plot = density_plots_diff[[plot_name]])
}
```

#Automatic whole range
```{r}
#Whole range
global_x_range <- range(unlist(lapply(BIS_columns_renamed, function(col) {
  diff_values <- Patients_pre_vs_post_imp[Patients_pre_vs_post_imp[["time"]] == "post", col] -
                 Patients_pre_vs_post_imp[Patients_pre_vs_post_imp[["time"]] == "pre", col]
  return(diff_values)
})), na.rm = TRUE)

create_density_plot_diff_global <- function(data, col, x_range) {
  # Calculate difference values (Post - Pre)
  diff_values <- data[data[["time"]] == "post", col] - data[data[["time"]] == "pre", col]
  
  ggplot(data.frame(x = diff_values), aes(x = x)) +
    geom_density(fill = "red", alpha = 0.5) +
    labs(title = paste("Difference:", col),
         x = paste("Difference in", col),
         y = "Density") +
    xlim(x_range[1], x_range[2])
}

density_plots_diff <- lapply(BIS_columns_renamed, function(col) {
  create_density_plot_diff_global(Patients_pre_vs_post_imp, col, global_x_range)
})

names(density_plots_diff) <- BIS_columns_renamed
for (plot in density_plots_diff) {
  print(plot)
}

save_directory <- "Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Density Plot_PrePost/Automatic whole range"
for (plot_name in names(density_plots_diff)) {
  file_name <- paste0(save_directory, "/", plot_name, ".png")
  ggsave(file_name, plot = density_plots_diff[[plot_name]])
}
```