---
title: "Preprocessing_for_ML"
author: "Rebecca Delfendahl"
date: "2024-08-06"
output: html_document
params:
  input_data_path: "Y://PsyThera//Projekte_Meinke//Old_projects//Labrotation_Rebecca//0_Datapreparation//Daten_Gruppenvergleich//new//RT_trimmed_RT_wrong_removed//BIS//outliers-removed"
---

```{r R Setup, include = FALSE}
source("../Useful_functions.R")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,rows.print = 6, cols.min.print = 3)
library(dplyr)
library(tidyverse)
library(caret)
```

# 0. Load relevant data

```{r}
basic_path <- params$input_data_path

task_data <- read.csv(file.path(basic_path,"Performance_scoresTask_Performance_Scores.csv")) # contains overall accuracy
data_pat_pre <- read.csv(file.path(basic_path,"Data_Patients_Pre.csv"))
```

# 1. Exclude patients due to missing values

```{r}
data_pat_pre_clean <- data_pat_pre %>%
  filter(!is.na(Response)) # equals to NA at T1_BAT_FAS_score and/or T3_BAT_FAS_score
```
`r nrow(data_pat_pre) - nrow(data_pat_pre_clean)` patients were removed due to missing FAS_T1 and/or FAS_T3 (i.e. response criterion).

# 2. Add overall accuracy as additional feature
Note: also add PercentRmv as control variable if trials with wrong responses were discarded

```{r}
features_df <- left_join(x = data_pat_pre_clean,
                         y = task_data %>% select(Subject, NumbLet_Average_PC,
                                                  Stroop_Average_PC,
                                                  TwoBack_Average_PC),
                         by = "Subject")
```

# 3. Add classes (for classifier) and outcome (for regressor)

## 3.1. Classes (Responder vs. Non-Responder)
```{r Save labels in separate dataframe}
labels_df <- features_df["Response"]
features_df$Response <- NULL
```

## 3.2. Regression: Outcome (Percent-change in FAS)
```{r}
outcome_df <- round(data.frame(FAS_perc_change = ((features_df$T3_BAT_FAS_score - features_df$T1_BAT_FAS_score) / features_df$T1_BAT_FAS_score) * 100) * (-1),2)
```

# 4. Dummy-Vars and renaming of features
The only variable that needed one-hot encoding was "Abschluss" (education). As we had no missings in this variable, we were allowed to perform one-hot encoding outside the loop. Otherwise, we would need to perform it inside the loop, just after imputation. 
As "Abschluss" (education) had two levels with a very low number of subjects (level 1/Basic education/Hauptschule: 1 subject, level 4/Other/Anderes: 4 subjects), we decided to turn it only into one dummy-coded variable, coding the presence of level 3/High education/Gymnasium 
```{r Prepare for ML}
# 0 Check missings in "Abschluss"
any(is.na(features_df$Abschluss))
# 0 Check frequency of levels
table(features_df$Abschluss)
# 1. OneHot-Encoding of variable "Abschluss"
features_df_enc <- features_df
features_df_enc$Abschluss_Gymnasium <- dplyr::recode(features_df_enc$Abschluss, "1" = 0, "2" = 0, "3" = 1, "4" = 1)


# 2. Drop variables "Abschluss", "Subject", "Gruppe", and "T3_BAT_FAS_score"
features_df_enc$Subject <- NULL
features_df_enc$Gruppe <- NULL
features_df_enc$Abschluss <- NULL
features_df_enc$T3_BAT_FAS_score <- NULL

# 3. Rename columns
features_df_enc <- features_df_enc %>%
  rename(is_woman = Geschlecht,
         Age = Alter,
         FAS = T1_BAT_FAS_score,
         BDI_II = T1_BAT_BDI_II_score,
         STAI_T = T1_BAT_STAI_T_score,
         BIS_11 = T1_BAT_BIS_11_score,
         Kirby_k = T1_BAT_Kirby_k_score,
         CFC_14 = T1_BAT_CFC_14_score,
         SRHI = T1_BAT_SRHI_score,
         Abschluss_Gymnasium = Abschluss_Gymnasium)
```

# 5. Recode missing values for binary and non-binary features

```{r Missing values}
#var_category <- find_bin_cols(features_df_enc_clean)
# TEST
bin_cols <- c("is_woman", "Abschluss_Gymnasium")
non_bin_cols <- setdiff(names(features_df_enc), bin_cols) 

for (feature in bin_cols) {
  features_df_enc[[feature]][is.na(features_df_enc[[feature]])] <- 77777
}
for (feature in non_bin_cols) {
  features_df_enc[[feature]][is.na(features_df_enc[[feature]])] <- 99999
}
```

# 6. Create different feature sets

```{r Reduced feature dataframe}
# Set 1: Clinical features only
clin_features <- c("is_woman", "Age", "FAS", "BDI_II", "STAI_T", "BIS_11", "Kirby_k", "CFC_14", "SRHI", "Abschluss_Gymnasium")
clin_features_df_enc <- features_df_enc[clin_features]

# Set 2: All features but without Overall Accuracy
all_features_final <- features_df_enc[, !names(features_df_enc) %in% c("NumbLet_Average_PC", "Stroop_Average_PC", "TwoBack_Average_PC")]

# Set 3: features_df_enc, includes Overall Accuracy
```

# 7. Save all 5 dataframes

```{r Save dataframes}
output_mainpath <- "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/2_Machine_learning/Feature_Label_Dataframes"
output_path <- create_results_folder(inputdata_path = params$input_data_path,output_mainpath = output_mainpath)

write.csv(all_features_final, file.path(output_path, "all_features.csv"), row.names = FALSE)
write.csv(features_df_enc, file.path(output_path, "all_features_accuracy.csv"), row.names = FALSE)
write.csv(clin_features_df_enc, file.path(output_path, "clinical_features_only.csv"), row.names = FALSE)
write.csv(labels_df, file.path(output_path, "labels.csv"), row.names = FALSE)
write.csv(outcome_df, file.path(output_path, "outcomes.csv"), row.names = FALSE)
```