---
title: "Descriptive tables (publication-ready): Healthy controls vs patients"
authors: "Rebecca Delfendahl, Till Julius Adam, Charlotte Meinke"
date: "2025-04-15"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
params:
  input_data_path: "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/0_Datapreparation/Taskdata/BIS"
  output_base_path: "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/1_Group_comparison/HC_vs_Pat"
---
<style type="text/css">
.main-container { /* Adjust main blocks */
  max-width: 100% !important;
  margin: auto;
}

body {
  font-family: "Georgia", serif !important; /* Set the font for the entire document */
}

.tocify { /* Adjust table of contents */
  max-width: 100% !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(gtsummary)
library(tidyr)
library(flextable)

# User-defined functions
source("../Useful_functions.R")
```

# Set-Up
## 1. Settings flextable
```{r}
format_flextable_portrait <- flextable_settings(word_orientation = "portrait")
format_flextable_landscape <- flextable_settings(word_orientation = "landscape")
```

## 2. Import Data (HCs and Patients_Pre)
```{r}
task_file_path <- file.path(params$input_data_path, "BIS_RT_trimmed_RT_wrong_removed_outliers-removed.csv")
task_data_all <- read.csv(task_file_path)

# Get Pat_Pre and HC data
Patients_vs_HC <- get_group_data(data = task_data_all, group = c("HC", "Pat_Pre"))
```

## 3. Set results path to save tables and images
```{r}
results_path_HC_Pat <- create_results_path(inputdata_path = basic_path, 
                                           output_mainpath = params$output_base_path)
```

## 4. Add socdem and clinical data
```{r}
socdem_file_path = "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/0_Datapreparation/Clinical_data/socdem-data_cleaned_response.csv"
socdem_clin_vars <- c("Alter", "Geschlecht", "Abschluss_Gymnasium", "T1_BAT_FAS_score", "T1_BAT_BDI_II_score", "T1_BAT_STAI_T_score", "T1_BAT_BIS_11_score", "T1_BAT_Kirby_k_score", "T1_BAT_CFC_14_score", "T1_BAT_SRHI_score", "BAT_T1")
Patients_vs_HC <- add_socdem_vars(df = Patients_vs_HC, socdem_file_path = socdem_file_path, vars = socdem_clin_vars)
```


# Descriptive group comparisons

## 0. Define categorical vs. dimensional variables
```{r}
categorical_vars <- c("Geschlecht", "Abschluss_Gymnasium")
dimensional_vars <- setdiff(socdem_clin_vars, categorical_vars)
```

## 1. Welch t-test for dimensional variables
```{r}
t_test_descr_HC_Pat <- t_test_mult_cols(df_basis = Patients_vs_HC, cols = dimensional_vars, grouping_variable = "Gruppe")
```

## 2. Chi-square test for categorical variables
```{r}
chi_sq_test_descr_HC_Pat <- chi_sq_test_mult_cols(df_basis = Patients_vs_HC, cols = categorical_vars, grouping_variable = "Gruppe")
```

## 3. Converge tables format and merge
```{r}
t_test_descr_pub_HC_Pat <- prepare_ttest_table(ttest_table = t_test_descr_HC_Pat, var_type = "dimensional", group_0 = "Healthy Controls", group_1 = "Patients")
chi_sq_test_descr_pub_HC_Pat <- prepare_ttest_table(ttest_table = chi_sq_test_descr_HC_Pat, var_type = "categorical", group_0 = "Healthy Controls", group_1 = "Patients")

# Merge the dataframes
descr_comps_pub_HC_Pat <- rbind(t_test_descr_pub_HC_Pat, chi_sq_test_descr_pub_HC_Pat)
# Reorder and rename the variables
descr_comps_pub_HC_Pat <- reorder_and_rename_rows(df = descr_comps_pub_HC_Pat, col_name = "Variable", label_map = variable_labels)

# Print missings
for (i in seq_len(nrow(descr_comps_pub_HC_Pat))) {
  miss_hc <- descr_comps_pub_HC_Pat$`Missings_Healthy Controls`[i]
  miss_pat <- descr_comps_pub_HC_Pat$Missings_Patients[i]
  
  if (miss_hc > 0 || miss_pat > 0) {
    var_name <- descr_comps_pub_HC_Pat$Variable[i]
    cat(sprintf("Missing values **%s**: Healthy Controls = %s, Patients = %s \n",
                var_name, miss_hc, miss_pat))
  }
}

# Remove columns Missings_HC, Missings_Patients
descr_comps_pub_HC_Pat <- descr_comps_pub_HC_Pat[, !(names(descr_comps_pub_HC_Pat) %in% c("Missings_Healthy Controls", "Missings_Patients"))]

# Add number of subjects per group
n_HC <- sum(Patients_vs_HC$Gruppe == 0)
n_Patients <- sum(Patients_vs_HC$Gruppe == 1)
n_subjects_row <- data.frame(Variable = "N", `Healthy Controls` = n_HC, Patients = n_Patients, Statistic = "-", stringsAsFactors = FALSE, check.names = FALSE)
descr_comps_pub_HC_Pat <- rbind(n_subjects_row, descr_comps_pub_HC_Pat)
```

## 3. Create and save flextable
```{r}
create_save_flextable(table_pub = descr_comps_pub_HC_Pat, results_path = results_path_HC_Pat, file_name = "descr_tests_Patients_HCs.docx")
```


# Descriptive Statistics

## 1. Prepare task-specific data frame of characteristics
```{r}
# 1. Subset data frame by non-NA for respective tasks
task_list <- list(
  "2-Back Task" = list(
    subset_data = Patients_vs_HC %>%
      filter(!is.na(TwoBack_BIS_Foil) & 
             !is.na(TwoBack_BIS_Target) & 
             !is.na(TwoBack_BIS_Total))
  ),
  "Stroop Task" = list(
    subset_data = Patients_vs_HC %>%
      filter(!is.na(Stroop_BIS_Congruent) & 
             !is.na(Stroop_BIS_Incongruent) & 
             !is.na(Stroop_BIS_Diff_Score))
  ),
  "Number-Letter Task" = list(
    subset_data = Patients_vs_HC %>%
      filter(!is.na(NumberLetter_BIS_Repeat) & 
             !is.na(NumberLetter_BIS_Switch) & 
             !is.na(NumberLetter_BIS_Diff_Score))
  ),
  "Stop-Signal Task" = list(
    subset_data = Patients_vs_HC %>%
      filter(!is.na(SSRT))
  )
)

# Loop through the tasks and create tables
characteristics_tables <- lapply(names(task_list), function(task_name) {
  data <- task_list[[task_name]]$subset_data
  generate_characteristics_table(data, task_name, grouping_variable = Gruppe, group_labels = c("Healthy Controls", "Patients"))
})

# Name the tables for easy reference
names(characteristics_tables) <- names(task_list)


# 3. Combined data frame of task-specific descriptive characteristics
# Calculate N for Healthy Controls and Patients for each task
calculate_n <- function(data) {
  data %>%
    group_by(Gruppe) %>%
    summarize(N = n(), .groups = "drop") %>%
    pivot_wider(names_from = Gruppe, values_from = N, names_prefix = "Group_") %>%
    rename("Healthy Controls" = Group_0, "Patients" = Group_1)
}

# Get N for each task
stroop_n <- calculate_n(task_list[["Stroop Task"]]$subset_data)
number_letter_n <- calculate_n(task_list[["Number-Letter Task"]]$subset_data)
two_back_n <- calculate_n(task_list[["2-Back Task"]]$subset_data)
ssrt_n <- calculate_n(task_list[["Stop-Signal Task"]]$subset_data)

# Convert the tbl_summary tables to tibbles and rename columns dynamically (no information on specific N needed)
stroop_df <- as_tibble(characteristics_tables[["Stroop Task"]]) %>%
  rename_with(~ c("Healthy Controls", "Patients"), .cols = 2:3)

number_letter_df <- as_tibble(characteristics_tables[["Number-Letter Task"]]) %>%
  rename_with(~ c("Healthy Controls", "Patients"), .cols = 2:3)

two_back_df <- as_tibble(characteristics_tables[["2-Back Task"]]) %>%
  rename_with(~ c("Healthy Controls", "Patients"), .cols = 2:3)

ssrt_df <- as_tibble(characteristics_tables[["Stop-Signal Task"]]) %>%
  rename_with(~ c("Healthy Controls", "Patients"), .cols = 2:3)

# Insert the "N" row into each data frame
stroop_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Healthy Controls" = stroop_n[["Healthy Controls"]], 
         "Patients" = stroop_n[["Patients"]]),
  stroop_df
)
number_letter_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Healthy Controls" = number_letter_n[["Healthy Controls"]], 
         "Patients" = number_letter_n[["Patients"]]),
  number_letter_df
)
two_back_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Healthy Controls" = two_back_n[["Healthy Controls"]], 
         "Patients" = two_back_n[["Patients"]]),
  two_back_df
)
ssrt_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Healthy Controls" = ssrt_n[["Healthy Controls"]], 
         "Patients" = ssrt_n[["Patients"]]),
  ssrt_df
)

# Combine the modified data frames
characteristics_tasks <- cbind(
  two_back_df[, "**Characteristic**", drop = FALSE],  # Include "Characteristic" column only once
  two_back_df[, -which(names(two_back_df) == "**Characteristic**")],
  stroop_df[, -which(names(stroop_df) == "**Characteristic**")],
  number_letter_df[, -which(names(number_letter_df) == "**Characteristic**")],
  ssrt_df[, -which(names(ssrt_df) == "**Characteristic**")]  # Drop "Characteristic" column from all others
)
```

## 2. Flextable of characteristics by task
```{r}
# Create Flextable
# Rename columns to ensure uniqueness
colnames(characteristics_tasks) <- c(
  "Characteristic", 
  "Two_Back_Healthy_Controls", "Two_Back_Patients", 
  "Stroop_Healthy_Controls", "Stroop_Patients",
  "Number_Letter_Healthy_Controls", "Number_Letter_Patients", 
  "Stop_Signal_Healthy_Controls", "Stop_Signal_Patients"
)

# Define primary (task) and secondary (HC vs Pat) headers
primary_headers <- c("Characteristic", "2-Back Task", "2-Back Task", 
                     "Stroop Task", "Stroop Task", 
                     "Number-Letter Task", "Number-Letter Task", 
                     "Stop-Signal Task", "Stop-Signal Task")

secondary_headers <- c("Characteristic", "Healthy Controls", "Patients", 
                       "Healthy Controls", "Patients", 
                       "Healthy Controls", "Patients", 
                       "Healthy Controls", "Patients")

# Create a flextable
ft <- flextable(characteristics_tasks)

# Add primary and secondary column headers
ft <- set_header_df(
  x = ft,
  mapping = data.frame(
    keys = colnames(characteristics_tasks),
    primary = primary_headers,
    secondary = secondary_headers,
    stringsAsFactors = FALSE
  ),
  key = "keys"
)

# Merge the primary headers to span across related columns
ft <- merge_h(ft, part = "header")

# Align and format the headers
ft <- align(ft, align = "center", part = "header")
ft <- bold(ft, part = "header")
ft <- autofit(ft)

# Write to Word
save_as_docx(
  ft,
  path =file.path(results_path_HC_Pat, "Characteristics_Patients_HCs.docx"), 
  pr_section = format_flextable_portrait)
ft
```

# Removed trials per task

```{r}
# Select columns
columns_n_trimmed <- grep("_n_trimmed_removed_RT$", names(Patients_vs_HC), value = TRUE)
columns_n_wrong <- grep("_n_wrong_removed_RT$", names(Patients_vs_HC), value = TRUE)

# Create a data frame dynamically
df <- data.frame(
  Condition = gsub("_n_trimmed_removed_RT", "", columns_n_trimmed),  # Remove suffix for clean names
  n_trimmed_removed = sapply(Patients_vs_HC[ columns_n_trimmed], mean, na.rm = TRUE),
  n_trimmed_removed_sd = sapply(Patients_vs_HC[ columns_n_trimmed], sd, na.rm = TRUE),
  n_wrong_removed = sapply(Patients_vs_HC[columns_n_wrong], mean, na.rm = TRUE),
  n_wrong_removed_sd = sapply(Patients_vs_HC[columns_n_wrong], sd, na.rm = TRUE)
)

# Rename variables
df$Condition <- dplyr::recode(df$Condition, Repeat = "Number-Letter: Repeat",
                          Switch = "Number-Letter: Switch", Congruent = "Stroop: Congruent",
                          Incongruent = "Stroop: Incongruent", Foil = "2-Back: Foil", 
                          Target = "2-Back: Target")
colnames(df) <- c("Condition", "Number of trials removed due to trimming (Mean)",
                  "Number of trials removed due to trimming (SD)",
                  "Number of wrong trials removed (Mean)",
                  "Number of wrong trials removed (SD)")

# Turn to flextable
ft_removedRT <- flextable(df)

# Round to 2
ft_removedRT <- colformat_double(ft_removedRT, digits = 2)

# Set table properties
ft_removedRT <- set_table_properties(ft_removedRT, width = 1, layout = "autofit")

# Header in bold
ft_removedRT <- bold(ft_removedRT, bold = TRUE, part = "header")

# Alignments
ft_removedRT<- align(ft_removedRT, j = 1, align = "left", part = "all") # first column
ft_removedRT <- align(ft_removedRT, j = 2:ncol(df), align = "center", part = "all") # rest

# Export flextable
save_as_docx(
  ft_removedRT,
  path = file.path(results_path_HC_Pat, "removed_trials_Patients_HCs.docx"), 
  pr_section = format_flextable_portrait)

# Info for interpreting the results: 
# Number-letter: switch 64, Repeat: 63
# Stroop: Congruent: 80 trials, Incongruent: 80 trials
# Spatial 2-back: Foil: 48 trials, Target: 112 trials
```
