---
title: "Group Comparison: Responders vs. Nonresponders"
authors: "Charlotte Meinke, Rebecca Delfendahl, Till Julius Adam"
date: "2024-09-04"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
params:
  input_data_path: "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/0_Datapreparation/Taskdata/BIS/BIS_RT-trimmed-RT-wrong-removed_outliers-removed.csv"
  output_base_path: "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/1_Group_comparison/R_vs_NR"
  response_criterion: "Response_FSQ_50" # choose between "Response_FSQ_50", "Response_FSQ_clinsig", and "Response_BAT"
  
---
<style type="text/css">
.main-container { /* Adjust main blocks */
  max-width: 100% !important;
  margin: auto;
}

body {
  font-family: "Georgia", serif !important; /* Set the font for the entire document */
}

.tocify { /* Adjust table of contents */
  max-width: 100% !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(pander)
library(effsize)
library(pwr)
library(gtsummary)
library(gt)
library(car)
library(flextable)
#library(officer)

# User-defined functions
source("../Useful_functions.R")
```

# Set-Up
## 1. Settings flextable
```{r}
format_flextable_portrait <- flextable_settings(word_orientation = "portrait")
format_flextable_landscape <- flextable_settings(word_orientation = "landscape")
```

## 2. Import Data 
```{r}
task_data_all <- read.csv(params$input_data_path)

# Get Pat_Pre data
data_Pat_pre <- get_group_data(data = task_data_all, group = c("Pat_Pre"))
```

## 3. Set results path to save tables and images
```{r}
results_path <- create_results_path(inputdata_path = params$input_data_path, 
                                         output_mainpath = params$output_base_path,
                                         response_criterion =  params$response_criterion)
```

## 4. Prepare dataset
### 4.1. Add response criterion
```{r}
socdem_file_path = "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/0_Datapreparation/Clinical_data/socdem-data_cleaned_response-added.csv"
data_Pat_pre <- add_response(df = data_Pat_pre, socdem_file_path = socdem_file_path, criterion = params$response_criterion)
```

### 4.2. Remove all patients that had no post-treatment score
```{r}
data_Pat_pre_clean <- data_Pat_pre[!is.na(data_Pat_pre$Response),]
```
We had `r nrow(data_Pat_pre[!is.na(data_Pat_pre$Response),])` patients with a post-value available, and `r nrow(data_Pat_pre_clean[data_Pat_pre_clean$Response == 1,])`


# Inferential statistics
### 1. Preparation
```{r}
# Ensure "Response" is treated as a factor
data_Pat_pre_clean$Response <- as.factor(data_Pat_pre_clean$Response)

# Define the columns for which to perform the test (i.e., task performance measures)
imp_columns <- c("NumberLetter_BIS_Repeat", "NumberLetter_BIS_Switch", 
                 "NumberLetter_BIS_Diff_Score", "Stroop_BIS_Congruent", 
                 "Stroop_BIS_Incongruent", "Stroop_BIS_Diff_Score", 
                 "TwoBack_BIS_Foil", "TwoBack_BIS_Target", "TwoBack_BIS_Total", 
                 "StopSignal_SSRT")
```

### 2. Test for variance homogeneity
This test is actually not necessary as we follow the recommendation 
```{r}
levene_test_table <- levene_test_mult_cols(df_basis = data_Pat_pre_clean, cols = imp_columns, grouping_variable = "Response")
#pander(levene_test_table, style = "rmarkdown", fontsize = "tiny")
levene_test_table
```
Levene tests show variance homogeneity for all task conditions between responders and non-responders.

### 3. Calculate Welch t-tests 
```{r}
t_test_table <- t_test_mult_cols(df_basis = data_Pat_pre_clean, cols = imp_columns, grouping_variable = "Response")
#pander(t_test_table, style = "rmarkdown", split.table = Inf, fontsize = "tiny")
t_test_table
```

### 4. Prepare table for publication
```{r table preparation, echo=FALSE, warning = FALSE}

t_test_table_pub <- cbind(performance_measures = rownames(t_test_table), t_test_table)

# Combine "Mean" and "SD" columns and reduce dataframe
t_test_table_pub <- t_test_table_pub %>%
  mutate(
    `Responder Mean (SD)` = paste0(group_1_mean, " (", group_1_sd, ")"),
    `Nonresponder Mean (SD)` = paste0(group_0_mean, " (", group_0_sd, ")")
    ) %>%
  select(performance_measures, `Responder Mean (SD)`, `Nonresponder Mean (SD)`, df, statistic, effect_size, p_value, p_value_adjusted)

# Change colnames
colnames(t_test_table_pub) <- c("Performance Measure", "Responder Mean (SD)", "Non-Responder Mean (SD)", "df", "t-Statistic", "Hedges' g","p-Value", "adj. p-Value")

# Rename performance measures using a named vector
t_test_table_pub <- reorder_and_rename_rows(df = t_test_table_pub, col_name = "Performance Measure", label_map = variable_labels)
```

### 5. Create and save flextable
```{r flextablesettings, echo=FALSE, warning = FALSE}
ft_t_test <- flextable::flextable(t_test_table_pub)

# Set table properties
ft_t_test <- flextable::set_table_properties(ft_t_test, width = 1, layout = "autofit")

# Header in bold
ft_t_test <- flextable::bold(ft_t_test, bold = TRUE, part = "header")

# Alignments
ft_t_test <- flextable::align(ft_t_test, j = 1, align = "left", part = "all") # first column
ft_t_test <- flextable::align(ft_t_test, j = 2:ncol(t_test_table_pub), align = "center", part = "all") # rest

# Export flextable
flextable::save_as_docx(
  ft_t_test,
  path = file.path(results_path, "t-test_Responder_Nonresponder.docx"), 
  pr_section = format_flextable_portrait)
```


# Plotting
### 1. Prepare data
```{r}
# Reshape data from wide to long format, including SSRT, for Response vs. Nonresponse
Response_vs_NonResponse_long <- data_Pat_pre_clean %>%
  pivot_longer(cols = all_of(imp_columns), # imp_columns now includes SSRT
               names_to = "Condition",
               values_to = "BIS_Score") %>%
  drop_na()

# Add a new task variable for faceting
Response_vs_NonResponse_long <- Response_vs_NonResponse_long %>%
  mutate(Task = case_when(
    grepl("NumberLetter", Condition) ~ "Number-Letter Task",
    grepl("Stroop", Condition) ~ "Stroop Task",
    grepl("TwoBack", Condition) ~ "2-Back Task",
    Condition == "SSRT" ~ "Stop Signal Task"
  ))

# Remove prefixes from Condition labels, keeping Task intact
Response_vs_NonResponse_long <- Response_vs_NonResponse_long %>%
  mutate(Condition = gsub("NumberLetter_BIS_|Stroop_BIS_|TwoBack_BIS_", "", Condition)) %>%
  mutate(Condition = as.character(Condition)) %>%
  
  # Set the condition order for each task separately to avoid conflicts (cannot both be "Diff_Score")
  mutate(Condition = ifelse(Task == "NumberLetter" & Condition == "Diff_Score", "Diff_Score NL", Condition)) %>%
  mutate(Condition = ifelse(Task == "Stroop" & Condition == "Diff_Score", "Diff_Score Stroop", Condition)) %>%
  
  # Convert Condition to factor with ordering
  mutate(Condition = case_when(
    Task == "NumberLetter" ~ factor(Condition, levels = c("Repeat", "Switch", "Diff_Score NL")),
    Task == "Stroop" ~ factor(Condition, levels = c("Congruent", "Incongruent", "Diff_Score Stroop")),
    Task == "TwoBack" ~ factor(Condition, levels = c("Target", "Foil", "Total")),
    TRUE ~ as.factor(Condition)
  ))

# Convert "Response" to factor
Response_vs_NonResponse_long$Response <- factor(Response_vs_NonResponse_long$Response, levels = c(0, 1), labels = c("Nonresponders", "Responders"))
```

### 2. Violin Plot
```{r, fig.width = 16, fig.height = 9, warning=FALSE}
# Create violin plots with mean and rename conditions in the plot

violin_plots_response_vs_nonresponse <- ggplot(Response_vs_NonResponse_long, aes(x = Condition, y = BIS_Score, fill = Response)) +
  geom_violin(position = position_dodge(width = 0.8)) +
  stat_summary(fun.data = mean_sdl, geom = "pointrange", # mult = 2 (SD) by default
               position = position_dodge(width = 0.8)) + # width = 0.65, fatten = 2
  facet_wrap(~ Task, scales = "free") +
  labs(x = "",
       y = "Performance Score",
       fill = "Group") +
  scale_x_discrete(labels = function(x) {
    x[x == "Diff_Score"] <- "Difference \nScore"
    x[x == "Diff_Score Stroop"] <- "Difference \nScore"
    return(x)
  }) +
  scale_fill_manual(values = c("Nonresponders" = "#D3D3D3", #colors 
                               "Responders" = "#48CFCB")) +
  theme_bw() +
  theme(
    text = element_text(family = "Arial"), #Font
    axis.title = element_text(size = 24),
    axis.title.y = element_text(size = 24, margin = margin(r = 30)),
    axis.text = element_text(size = 18),
    legend.title = element_text(size = 24),
    legend.text = element_text(size = 18),
    strip.text = element_text(size = 20),
    panel.spacing.x = unit(3, "lines")
  )
#print(violin_plots_response_vs_nonresponse)

ggsave(file.path(results_path,"violin_plots_response_nonresponse.png"), 
       plot = violin_plots_response_vs_nonresponse, width = 16, height = 9, dpi = 300)
ggsave(file.path(results_path, "violin_plots_response_nonresponse.svg"), 
       plot = violin_plots_response_vs_nonresponse, width = 16, height = 9, dpi = 300)

```
