---
title: "Descriptive tables (publication-ready): Responder vs non-responder"
author: "Rebecca Delfendahl"
date: "2025-05-14"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
params:
  input_data_path: "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/0_Datapreparation/RT_trimmed_RT_wrong_removed/BIS/outliers-removed"
  output_base_path: "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/1_Group_comparison/R_vs_NR"
  response_criterion: "response_FSQ" # choose between "response_FSQ" or "response_BAT"
---
<style type="text/css">
.main-container { /* Adjust main blocks */
  max-width: 100% !important;
  margin: auto;
}

body {
  font-family: "Georgia", serif !important; /* Set the font for the entire document */
}

.tocify { /* Adjust table of contents */
  max-width: 100% !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(gtsummary)
library(tidyr)
library(flextable)
library(officer)

# User-defined functions
source("../Useful_functions.R")
```

# Set-Up
## 1. Load settings flextable
```{r}
format_flextable_portrait <- flextable_settings(word_orientation = "portrait")
format_flextable_landscape <- flextable_settings(word_orientation = "landscape")
```

## 2. Import Data (Patients_Pre)
```{r}
basic_path <- file.path(params$input_data_path, params$response_criterion)
data_Pat_pre <- read.csv(file.path(basic_path,"Data_Patients_Pre.csv"))
```

## 3. Set results path to save tables and images
```{r}
results_path <- create_results_path(inputdata_path = basic_path, 
                                    output_mainpath = params$output_base_path,
                                    response_criterion =  params$response_criterion)
```

## 4. Prepare dataset
Remove all patients that had no post-treatment score
```{r}
data_Pat_pre_clean <- data_Pat_pre[!is.na(data_Pat_pre$Response),]

# Dummy-code variable education
recode_results_2 <- dummy_code_education(data_Pat_pre_clean)
data_Pat_pre_clean <- recode_results_2$data_recoded
```
We had `r nrow(data_Pat_pre[!is.na(data_Pat_pre$Response),])` patients with a post-value available, and `r nrow(data_Pat_pre_clean[data_Pat_pre_clean$Response == 1,])`


# Descriptive group comparisons

## 0. Define categorical vs. dimensional variables
```{r}
categorical_vars <- c("Geschlecht", "Abschluss_Gymnasium")
dimensional_vars <- c("Alter", "T1_BAT_FAS_score", "T1_BAT_BDI_II_score", "T1_BAT_STAI_T_score", "T1_BAT_BIS_11_score", "T1_BAT_Kirby_k_score", "T1_BAT_CFC_14_score", "T1_BAT_SRHI_score", "BAT_T1")
```

## 1. Welch t-test for dimensional variables
```{r}
t_test_descr_Response <- t_test_mult_cols(df_basis = data_Pat_pre_clean, cols = dimensional_vars, grouping_variable = "Response")
```

## 2. Chi-square test for categorical variables
```{r}
chi_sq_test_descr_Response <- chi_sq_test_mult_cols(df_basis = data_Pat_pre_clean, cols = categorical_vars, grouping_variable = "Response")
```

## 3. Converge tables format and merge
```{r}
t_test_descr_pub_Response <- prepare_ttest_table(ttest_table = t_test_descr_Response, var_type = "dimensional", group_0 = "Non-Responder", group_1 = "Responder")
chi_sq_test_descr_pub_Response <- prepare_ttest_table(ttest_table = chi_sq_test_descr_Response, var_type = "categorical", group_0 = "Non-Responder", group_1 = "Responder")

# Merge the dataframes
descr_comps_Response <- rbind(t_test_descr_pub_Response, chi_sq_test_descr_pub_Response)
# Reorder and rename the variables
descr_comps_Response <- reorder_and_rename_rows(df = descr_comps_Response, col_name = "Variable", label_map = variable_labels)

# Print missings
for (i in seq_len(nrow(descr_comps_Response))) {
  miss_NR <- descr_comps_Response$`Missings_Non-Responder`[i]
  miss_R <- descr_comps_Response$Missings_Responder[i]
  
  if (miss_NR > 0 || miss_R > 0) {
    var_name <- descr_comps_Response$Variable[i]
    cat(sprintf("Missing values **%s**: Non-Responder = %s, Responder = %s \n",
                var_name, miss_NR, miss_R))
  }
}

# Remove columns Missings_HC, Missings_Patients
descr_comps_Response <- descr_comps_Response[, !(names(descr_comps_Response) %in% c("Missings_Non-Responder", "Missings_Responder"))]

# Add number of subjects per group
n_NR <- sum(data_Pat_pre_clean$Response == 0)
n_R <- sum(data_Pat_pre_clean$Response == 1)
n_subjects_row <- data.frame(Variable = "N", `Non-Responder` = n_NR, Responder = n_R, Statistic = "-", stringsAsFactors = FALSE, check.names = FALSE)
descr_comps_Response <- rbind(n_subjects_row, descr_comps_Response)
```

## 3. Create and save flextable
```{r}
create_save_flextable(table_pub = descr_comps_Response, results_path = results_path, file_name = "descr_tests_R_vs_NR.docx")
```


# Descriptive Statistics

## 1. Prepare task-specific data frame of characteristics
```{r}
# 1. Subset data frame by non-NA for respective tasks
task_list <- list(
  "Stroop Task" = list(
    subset_data = data_Pat_pre_clean %>%
      filter(!is.na(Stroop_BIS_Congruent) & 
             !is.na(Stroop_BIS_Incongruent) & 
             !is.na(Stroop_BIS_Diff_Score))
  ),
  "Number-Letter Task" = list(
    subset_data = data_Pat_pre_clean %>%
      filter(!is.na(NumberLetter_BIS_Repeat) & 
             !is.na(NumberLetter_BIS_Switch) & 
             !is.na(NumberLetter_BIS_Diff_Score))
  ),
  "2-Back Task" = list(
    subset_data = data_Pat_pre_clean %>%
      filter(!is.na(TwoBack_BIS_Foil) & 
             !is.na(TwoBack_BIS_Target) & 
             !is.na(TwoBack_BIS_Total))
  ),
  "Stop-Signal Task" = list(
    subset_data = data_Pat_pre_clean %>%
      filter(!is.na(SSRT))
  )
)

# Loop through the tasks and create tables
characteristics_tables <- lapply(names(task_list), function(task_name) {
  data <- task_list[[task_name]]$subset_data
  generate_characteristics_table(data, task_name, grouping_variable = Response, group_labels = c("Non-Response", "Response"))
})

# Name the tables for easy reference
names(characteristics_tables) <- names(task_list)


# 3. Combined data frame of task-specific descriptive characteristics
# Calculate N for Healthy Controls and Patients for each task
calculate_n <- function(data) {
  data %>%
    group_by(Response) %>%
    summarize(N = n(), .groups = "drop") %>%
    pivot_wider(names_from = Response, values_from = N, names_prefix = "Group_") %>%
    rename("Non-Response" = Group_0, "Response" = Group_1)
}

# Get N for each task
stroop_n <- calculate_n(task_list[["Stroop Task"]]$subset_data)
number_letter_n <- calculate_n(task_list[["Number-Letter Task"]]$subset_data)
two_back_n <- calculate_n(task_list[["2-Back Task"]]$subset_data)
ssrt_n <- calculate_n(task_list[["Stop-Signal Task"]]$subset_data)

# Convert the tbl_summary tables to tibbles and rename columns dynamically (no information on specific N needed)
stroop_df <- as_tibble(characteristics_tables[["Stroop Task"]]) %>%
  rename_with(~ c("Non-Response", "Response"), .cols = 2:3)

number_letter_df <- as_tibble(characteristics_tables[["Number-Letter Task"]]) %>%
  rename_with(~ c("Non-Response", "Response"), .cols = 2:3)

two_back_df <- as_tibble(characteristics_tables[["2-Back Task"]]) %>%
  rename_with(~ c("Non-Response", "Response"), .cols = 2:3)

ssrt_df <- as_tibble(characteristics_tables[["Stop-Signal Task"]]) %>%
  rename_with(~ c("Non-Response", "Response"), .cols = 2:3)

# Insert the "N" row into each data frame
stroop_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Non-Response" = stroop_n[["Non-Response"]], 
         "Response" = stroop_n[["Response"]]),
  stroop_df
)
number_letter_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Non-Response" = number_letter_n[["Non-Response"]], 
         "Response" = number_letter_n[["Response"]]),
  number_letter_df
)
two_back_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Non-Response" = two_back_n[["Non-Response"]], 
         "Response" = two_back_n[["Response"]]),
  two_back_df
)
ssrt_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Non-Response" = ssrt_n[["Non-Response"]], 
         "Response" = ssrt_n[["Response"]]),
  ssrt_df
)

# Combine the modified data frames
characteristics_tasks <- cbind(
  stroop_df[, "**Characteristic**", drop = FALSE],  # Include "Characteristic" column only once
  stroop_df[, -which(names(stroop_df) == "**Characteristic**")],
  number_letter_df[, -which(names(number_letter_df) == "**Characteristic**")],
  two_back_df[, -which(names(two_back_df) == "**Characteristic**")],
  ssrt_df[, -which(names(ssrt_df) == "**Characteristic**")]  # Drop "Characteristic" column from all others
)
```

## 2. Flextable of characteristics by task
```{r}
# Rename columns to ensure uniqueness
colnames(characteristics_tasks) <- c(
  "Characteristic", 
  "Stroop_NR", "Stroop_R", 
  "Number_Letter_NR", "Number_Letter_R", 
  "Two_Back_NR", "Two_Back_R", 
  "Stop_Signal_NR", "Stop_Signal_R"
)

# Define primary (task) and secondary (HC vs Pat) headers
primary_headers <- c("Characteristic", "Stroop Task", "Stroop Task", 
                     "Number-Letter Task", "Number-Letter Task", 
                     "2-Back Task", "2-Back Task", 
                     "Stop-Signal Task", "Stop-Signal Task")

secondary_headers <- c("Characteristic", "Non-Response", "Response", 
                       "Non-Response", "Response", 
                       "Non-Response", "Response", 
                       "Non-Response", "Response")

# Create a flextable
ft <- flextable(characteristics_tasks)

# Add primary and secondary column headers
ft <- set_header_df(
  x = ft,
  mapping = data.frame(
    keys = colnames(characteristics_tasks),
    primary = primary_headers,
    secondary = secondary_headers,
    stringsAsFactors = FALSE
  ),
  key = "keys"
)

# Merge the primary headers to span across related columns
ft <- merge_h(ft, part = "header")

# Align and format the headers
ft <- align(ft, align = "center", part = "header")
ft <- bold(ft, part = "header")
ft <- autofit(ft)

# Write to Word
save_as_docx(
  ft,
  path = file.path(results_path, "Characteristics_Response_Nonresponse.docx"), 
  pr_section = format_flextable_landscape)
ft
```

# Removed trials per task

```{r}
# Select columns
columns_n_trimmed <- grep("_n_trimmed_removed_RT$", names(data_Pat_pre_clean), value = TRUE)
columns_n_wrong <- grep("_n_wrong_removed_RT$", names(data_Pat_pre_clean), value = TRUE)

# Create a data frame dynamically
df <- data.frame(
  Condition = gsub("_n_trimmed_removed_RT", "", columns_n_trimmed),  # Remove suffix for clean names
  n_trimmed_removed = sapply(data_Pat_pre_clean[ columns_n_trimmed], mean, na.rm = TRUE),
  n_trimmed_removed_sd = sapply(data_Pat_pre_clean[ columns_n_trimmed], sd, na.rm = TRUE),
  n_wrong_removed = sapply(data_Pat_pre_clean[columns_n_wrong], mean, na.rm = TRUE),
  n_wrong_removed_sd = sapply(data_Pat_pre_clean[columns_n_wrong], sd, na.rm = TRUE)
)

# Rename variables
df$Condition <- dplyr::recode(df$Condition, Repeat = "Number-Letter: Repeat",
                          Switch = "Number-Letter: Switch", Congruent = "Stroop: Congruent",
                          Incongruent = "Stroop: Incongruent", Foil = "2-Back: Foil", 
                          Target = "2-Back: Target")
colnames(df) <- c("Condition", "Number of trials removed due to trimming (Mean)",
                  "Number of trials removed due to trimming (SD)",
                  "Number of wrong trials removed (Mean)",
                  "Number of wrong trials removed (SD)")

# Turn to flextable
ft_removedRT <- flextable(df)

# Round to 2
ft_removedRT <- colformat_double(ft_removedRT, digits = 2)

# Set table properties
ft_removedRT <- set_table_properties(ft_removedRT, width = 1, layout = "autofit")

# Header in bold
ft_removedRT <- bold(ft_removedRT, bold = TRUE, part = "header")

# Alignments
ft_removedRT<- align(ft_removedRT, j = 1, align = "left", part = "all") # first column
ft_removedRT <- align(ft_removedRT, j = 2:ncol(df), align = "center", part = "all") # rest

# Export flextable
save_as_docx(
  ft_removedRT,
  path = file.path(results_path, "removed_trials_Responder_Nonresponder.docx"), 
  pr_section = format_flextable_portrait)

# Info for interpreting the results: 
# Number-letter: switch 64, Repeat: 63
# Stroop: Congruent: 80 trials, Incongruent: 80 trials
# Spatial 2-back: Foil: 48 trials, Target: 112 trials
```