---
title: "Group Comparison: Healthy Controls vs. Patients"
authors: "Charlotte Meinke, Rebecca Delfendahl, Till Julius Adam"
date: "2024-09-04"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
params:
  input_data_path: "Y:\\PsyThera\\Projekte_Meinke\\Old_projects\\Labrotation_Rebecca\\0_Datapreparation\\Daten_Gruppenvergleich\\new\\RT_trimmed_RT_wrong_removed\\BIS\\outliers-removed"
---
<style type="text/css">
.main-container { /* Adjust main blocks */
  max-width: 100% !important;
  margin: auto;
}

body {
  font-family: "Georgia", serif !important; /* Set the font for the entire document */
}

.tocify { /* Adjust table of contents */
  max-width: 100% !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(pander)
library(effsize)
library(pwr)
library(gtsummary)
library(gt)
library(car)
```

# Importing Data
```{r}
basic_path = params$input_data_path
data_HC <- read.csv(file.path(basic_path,"Data_HC.csv"))
data_Pat_pre <- read.csv(file.path(basic_path,"Data_Patients_Pre.csv"))
```

# Prepare dataset
```{r}
## 1. Merge patients pre and controls
Patients_vs_HC <- rbind(data_Pat_pre, data_HC)

## 2. Reduce dataset to executive functions of interest (BIS columns and SSRT)
BIS_columns <- colnames(Patients_vs_HC)[grep("BIS", colnames(Patients_vs_HC))]
BIS_columns <- BIS_columns[BIS_columns != "T1_BAT_BIS_11_score"] # BIS-11 is a questionnaire
imp_columns <- c(BIS_columns,"SSRT")
Patients_vs_HC_imp <- Patients_vs_HC[,c("Subject","Gruppe",imp_columns)]

## 3. Add confounds (Alter, Geschlecht, Abschluss)
Patients_vs_HC_imp_conf <- merge(Patients_vs_HC[,c("Subject","Alter","Geschlecht","Abschluss")], Patients_vs_HC_imp, by = "Subject")
```
In total, data for `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1,])` patients and `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0,])` healthy controls were available. When outliers were removed, the number of subjects varied slightly: 
Number-Letter Task: `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0 & !is.na(Patients_vs_HC_imp$NumberLetter_BIS_Repeat),])` healthy controls,  `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1 & !is.na(Patients_vs_HC_imp$NumberLetter_BIS_Repeat),])`
Stroop Task: `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0 & !is.na(Patients_vs_HC_imp$Stroop_BIS_Congruent),])` healthy controls,  `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1 & !is.na(Patients_vs_HC_imp$Stroop_BIS_Congruent),])`
2back-Task: `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0 & !is.na(Patients_vs_HC_imp$TwoBack_BIS_Total),])` healthy controls,  `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1 & !is.na(Patients_vs_HC_imp$TwoBack_BIS_Total),])`
SSRT: `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0 & !is.na(Patients_vs_HC_imp$SSRT),])` healthy controls,  `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1 & !is.na(Patients_vs_HC_imp$SSRT),])`
Find assumption test for normality in script "Testing Normality".

# Inferential statistics
## Check variance homogeneity
```{r}
# Ensure "Gruppe" is treated as a factor
Patients_vs_HC_imp$Gruppe <- as.factor(Patients_vs_HC_imp$Gruppe)

# Function to perform Levene's test on multiple columns
levene_test_mult_cols <- function(df_basis, cols) {
  df <- data.frame(p_value = numeric(length(cols)))
  rownames(df) <- cols
  
  for (i in seq_along(cols)) {
    col <- cols[i]
    # Perform Levene's test, treating 'Gruppe' as a factor
    levene_result <- car::leveneTest(df_basis[[col]] ~ df_basis[["Gruppe"]])
    df[col, "p_value"] <- round(levene_result[1, "Pr(>F)"], 4)
  }
  
  return(df)
}

# Perform Levene's test for homogeneity of variance
levene_test_table <- levene_test_mult_cols(df_basis = Patients_vs_HC_imp, cols = imp_columns)
pander(levene_test_table, style = "rmarkdown", fontsize = "tiny")
levene_test_table
```

## Calculate Welch t-test
```{r}
t_test_mult_cols <- function(df_basis, cols) {
  # Calculate a t-test for multiple comparisons and store the results in a dataframe
  df <- data.frame(group_mean_Patients = numeric(length(cols)), sd_Patients = numeric(length(cols)),
                   group_mean_HC = numeric(length(cols)), sd_HC = numeric(length(cols)),df = numeric(length(cols)), t_statistic = numeric(length(cols)), p_value = numeric(length(cols)),
                   cohen_d = numeric(length(cols)))
  rownames(df) <- cols
  p_values_raw <- numeric(length(cols))  # Vector to store raw p-values
  
  for (i in seq_along(cols)) {
    col <- cols[i] #seq_along for sequencing column indices
    
    # Using the non-formula interface
    group0 <- na.omit(df_basis[df_basis[["Gruppe"]] == 0, col])
    group1 <- na.omit(df_basis[df_basis[["Gruppe"]] == 1, col])
    results <- t.test(group0, group1, paired = FALSE, var.equal = FALSE)
    #Alternativ using formula method:
      #results <- t.test(df_basis[[col]] ~ df_basis[["Gruppe"]], paired = FALSE, var.equal = FALSE)
    
    # Store raw p-values for BH correction
    p_values_raw[i] <- results$p.value
    
    # Calculate Cohen's d using the cohen.d function (effect size)
    cohen_d_result <- cohen.d(group0, group1, hedges.correction = FALSE)
    cohen_d <- cohen_d_result$estimate
    
    # Calculate power using pwr.t.test
    n0 <- length(group0)
    n1 <- length(group1)
    power_result <- pwr.t.test(d = cohen_d, n = min(n0, n1), sig.level = 0.05, type = "two.sample", alternative = "greater")
    power <- power_result$power
    
    # Store the results
    df[col, "t-statistic"] <- round(results$statistic, 2)
    df[col, "df"] <- round(results[["parameter"]][["df"]], 2)
    df[col, "p_value"] <- round(results$p.value, 2)
    df[col, "group_mean_Patients"] <- round(mean(group1), 2)
    df[col, "sd_Patients"] <- round(sd(group1), 2)
    df[col, "group_mean_HC"] <- round(mean(group0), 2)
    df[col, "sd_HC"] <- round(sd(group0), 2)
    df[col, "cohen_d"] <- round(cohen_d, 2)
    df[col, "power"] <- round(power, 2)
  }
  
  # Adjust p-values using Benjamini-Hochberg method for multiple testing of related tasks
  p_values_adjusted <- p.adjust(p_values_raw, method = "BH")
  df$p_value_adjusted <- round(p_values_adjusted, 2)
  
  return(df)
}

t_test_table <- t_test_mult_cols(df_basis = Patients_vs_HC_imp, cols = imp_columns)
#pander(t_test_table, style = "rmarkdown", fontsize = "tiny")
t_test_table
```

### Characteristics Table: Descriptive Statistics
```{r}
# Data frame that codes variables according to study protocol
characteristics_data <- Patients_vs_HC %>%
  select(Gruppe, Alter, Geschlecht, Abschluss, 
         T1_BAT_BDI_II_score, T1_BAT_BIS_11_score, T1_BAT_CFC_14_score, 
         T1_BAT_FAS_score, T1_BAT_Kirby_k_score, T1_BAT_SRHI_score, T1_BAT_STAI_T_score) %>%
  mutate(Gruppe = factor(Gruppe, levels = c(0, 1), labels = c("Healthy Controls", "Patients")),
         Geschlecht = factor(Geschlecht, levels = c(0, 1), labels = c("Male", "Female")),
         Abschluss = factor(Abschluss, levels = c(1, 2, 3, 4), 
                            labels = c("Basic secondary education", "Intermediate secondary education",
                                       "Higher secondary education", "Other")))

# Create a summary table with a subheading for "Baseline Assessment"
characteristics_table <- characteristics_data %>%
  tbl_summary(by = Gruppe,  # Compare by Gruppe (HC vs. Pat)
              label = list(
                Alter ~ "Age",
                Geschlecht ~ "Sex",
                Abschluss ~ "Education",
                T1_BAT_BDI_II_score ~ "BDI-II Score",
                T1_BAT_BIS_11_score ~ "BIS-11 Score",
                T1_BAT_CFC_14_score ~ "CFC-14 Score",
                T1_BAT_FAS_score ~ "FSQ Score",
                T1_BAT_Kirby_k_score ~ "Kirby k Score",
                T1_BAT_SRHI_score ~ "SRHI Score",
                T1_BAT_STAI_T_score ~ "STAI-T Score"
              ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n}",
      Geschlecht ~ "{n} ({p}%)"  # Show counts and percentages for Geschlecht
    ),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Characteristic**") %>%
  modify_spanning_header(
    starts_with("T1_BAT_") ~ "**Baseline Assessment**"
  ) %>%
  # Remove the row for "Male" in the "Geschlecht" variable
  modify_table_body(
    ~ .x %>% filter(!(variable == "Geschlecht" & label == "Male"))
  )

characteristics_table
```

```{r}
leveneTest(T1_BAT_STAI_T_score ~ as.factor(Gruppe), data = Patients_vs_HC)
t_test_result <- t.test(T1_BAT_STAI_T_score ~ Gruppe, data = Patients_vs_HC, var.equal = FALSE)
print(t_test_result)
```


###  Table: Inferential Statistics
```{r}
# Create a data frame with inferential statistics (p-value, adjusted p-value, Cohen's d, power)
inferential_stats_df <- data.frame(
  Task = c("Number Letter Repeat", "Number Letter Switch", "Number Letter Difference BIS Score",
           "Stroop Congruent", "Stroop Incongruent", "Stroop BIS Difference Score",
           "Two Back Foil", "Two Back Target", "Two Back Difference BIS Score", "SSRT"),  # Include the missing task
  unadjusted_p_value = t_test_table$p_value,
  adjusted_p_value = t_test_table$p_value_adjusted,
  cohen_d = t_test_table$cohen_d,
  power = t_test_table$power
)

# Create the inferential statistics table using gt
inferential_stats_table <- inferential_stats_df %>%
  gt() %>% # uses gt instead of tbl_summary because easier implementation for inferential tables for which data already computed somewhere else
  tab_header(
    title = "Inferential Statistics: Healthy Controls vs. Patients"
  ) %>%
  cols_label(
    Task = html("<b>Task</b>"),
    unadjusted_p_value = html("<b>Unadjusted p-value</b>"),
    adjusted_p_value = html("<b>Adjusted p-value</b>"),
    cohen_d = html("<b>Cohen's d</b>"),
    power = html("<b>Power</b>")
  ) %>%
  fmt_number(
    columns = c(unadjusted_p_value, adjusted_p_value, cohen_d, power),
    decimals = 2
  )

inferential_stats_table
```


## <u>Plotting</u>

#### Prepare data
```{r}
# Reshape data from wide to long format, including SSRT
Patients_vs_HC_long <- Patients_vs_HC_imp %>%
  pivot_longer(cols = all_of(imp_columns), # imp_columns now includes SSRT
               names_to = "Condition",
               values_to = "BIS_Score")

# Add new task variable to keep the headers intact for faceting
Patients_vs_HC_long <- Patients_vs_HC_long %>%
  mutate(Task = case_when(
    grepl("NumberLetter", Condition) ~ "Number-Letter Task",
    grepl("Stroop", Condition) ~ "Stroop Task",
    grepl("TwoBack", Condition) ~ "2-Back Task",
    Condition == "SSRT" ~ "Stop Signal Task"
  ))

# Remove prefixes from Condition labels, keeping Task intact
Patients_vs_HC_long <- Patients_vs_HC_long %>%
  mutate(Condition = gsub("NumberLetter_BIS_|Stroop_BIS_|TwoBack_BIS_", "", Condition)) %>%
  mutate(Condition = as.character(Condition)) %>%
  
  # Set the condition order for each task separately to avoid conflicts (cannot both be "Diff_Score")
  mutate(Condition = ifelse(Task == "NumberLetter" & Condition == "Diff_Score", "Diff_Score NL", Condition)) %>%
  mutate(Condition = ifelse(Task == "Stroop" & Condition == "Diff_Score", "Diff_Score Stroop", Condition)) %>%
  
  # Convert Condition to factor with ordering
  mutate(Condition = case_when(
    Task == "NumberLetter" ~ factor(Condition, levels = c("Repeat", "Switch", "Diff_Score NL")),
    Task == "Stroop" ~ factor(Condition, levels = c("Congruent", "Incongruent", "Diff_Score Stroop")),
    Task == "TwoBack" ~ factor(Condition, levels = c("Foil", "Target", "Total")),
    TRUE ~ as.factor(Condition)
  ))

# Convert "Gruppe" to factor
Patients_vs_HC_long$Gruppe <- factor(Patients_vs_HC_long$Gruppe, levels = c(0, 1), labels = c("Healthy Controls", "Patients"))
```

##### Violin Plot: Patients vs. Healthy Controls
```{r}
# Create violin plots with mean and rename conditions in the plot

violin_plots_patients_hc <- ggplot(Patients_vs_HC_long, aes(x = Condition, y = BIS_Score, fill = Gruppe)) +
  geom_violin(position = position_dodge(width = 0.8)) +
  stat_summary(fun.data = mean_sdl, geom = "pointrange", # mult = 2 (SD) by default
               position = position_dodge(width = 0.8)) + # width = 0.65, fatten = 2
  facet_wrap(~ Task, scales = "free") +
  labs(x = "",
       y = "Performance Score",
       fill = "Group") +
  scale_x_discrete(labels = function(x) {
    x[x == "Diff_Score"] <- "Difference \nScore"
    x[x == "Diff_Score Stroop"] <- "Difference \nScore"
    x[x == "Total"] <- "Difference \nScore"
    return(x)
  }) +
  scale_fill_manual(values = c("Healthy Controls" = "#D3D3D3", #colors 
                                 "Patients" = "#48CFCB")) +
  theme_bw() +
  theme(
    text = element_text(family = "Arial"), #Font
    axis.title = element_text(size = 24),
    axis.title.y = element_text(size = 24, margin = margin(r = 30)),
    axis.text = element_text(size = 18),
    legend.title = element_text(size = 24),
    legend.text = element_text(size = 18),
    strip.text = element_text(size = 20),
    panel.spacing.x = unit(3, "lines")
  )
print(violin_plots_patients_hc)
ggsave("violin_plots_patients_hc.png", plot = violin_plots_patients_hc, width = 16, height = 9, dpi = 300)
```

##### Boxplot: Patients vs. Healthy Controls
```{r}
# Create boxplots
boxplots_Patients_HC <- ggplot(Patients_vs_HC_long, aes(x = Condition, y = BIS_Score, color = Gruppe)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(position = position_jitterdodge(),
              size = 0.7,
              shape = 21,
              fill = NA) +
  facet_wrap(~ Task, scales = "free") +
  #facet_grid(rows = vars(Task), scales = "free_x") +
  labs(title = "Boxplots of Task Condition by Group",
       x = "Condition",
       y = "BIS-Score",
       color = "Group")
print(boxplots_Patients_HC)
```
