---
title: "Group Comparison: Responders vs. Nonresponders"
authors: "Charlotte Meinke, Rebecca Delfendahl, Till Julius Adam"
date: "2024-09-04"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
params:
  basic_path: "Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/0_Datapreparation/Daten_Gruppenvergleich/new/RT_trimmed_RT_wrong_removed/BIS/outliers-removed"
---
<style type="text/css">
.main-container { /* Adjust main blocks */
  max-width: 100% !important;
  margin: auto;
}

body {
  font-family: "Georgia", serif !important; /* Set the font for the entire document */
}

.tocify { /* Adjust table of contents */
  max-width: 100% !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(pander)
library(effsize)
library(pwr)
library(gtsummary)
library(gt)
library(car)
library(flextable)
library(officer)

# User-defined functions
source("../Useful_functions.R")
```

# Settings flextable
```{r, echo = FALSE}
format_flextable_portrait <- flextable_settings(word_orientation = "portrait")
format_flextable_landscape <- flextable_settings(word_orientation = "landscape")
```

# Import Data 

```{r}
basic_path = "Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/0_Datapreparation/Daten_Gruppenvergleich/With_wrong_responses/outliers-removed"
param_suffix = basename(basic_path)
data_Pat_pre <- read.csv(file.path(basic_path,"Data_Patients_Pre.csv"))
```

# Prepare dataset
Remove all patients that had no post-treatment score
```{r}
data_Pat_pre_clean <- data_Pat_pre[!is.na(data_Pat_pre$Response),]
```

# Descriptive Statistics
Characteristics Table
Task-specific characteristics tables using tbl_summary
```{r}
# 1. Subset data frame by non-NA for respective tasks
task_list <- list(
  "Stroop Task" = list(
    subset_data = data_Pat_pre_clean %>%
      filter(!is.na(Stroop_BIS_Congruent) & 
             !is.na(Stroop_BIS_Incongruent) & 
             !is.na(Stroop_BIS_Diff_Score))
  ),
  "Number-Letter Task" = list(
    subset_data = data_Pat_pre_clean %>%
      filter(!is.na(NumberLetter_BIS_Repeat) & 
             !is.na(NumberLetter_BIS_Switch) & 
             !is.na(NumberLetter_BIS_Diff_Score))
  ),
  "2-Back Task" = list(
    subset_data = data_Pat_pre_clean %>%
      filter(!is.na(TwoBack_BIS_Foil) & 
             !is.na(TwoBack_BIS_Target) & 
             !is.na(TwoBack_BIS_Total))
  ),
  "Stop Signal Task" = list(
    subset_data = data_Pat_pre_clean %>%
      filter(!is.na(SSRT))
  )
)
```

```{r}
# 2. Function to generate characteristics table
generate_characteristics_table <- function(data, task_name) {
  data %>%
    select(Response, Alter, Geschlecht, Abschluss, 
           T1_BAT_BDI_II_score, T1_BAT_BIS_11_score, T1_BAT_CFC_14_score, 
           T1_BAT_FAS_score, T1_BAT_Kirby_k_score, T1_BAT_SRHI_score, T1_BAT_STAI_T_score) %>%
    mutate(Response = factor(Response, levels = c(1, 0), labels = c("Response", "Non-Response")),
           Geschlecht = factor(Geschlecht, levels = c(0, 1), labels = c("Male", "Female")),
           Abschluss = factor(Abschluss, levels = c(1, 2, 3, 4), 
                              labels = c("Basic secondary education", 
                                         "Intermediate secondary education",
                                         "Higher secondary education", 
                                         "Other"))) %>%
    tbl_summary(by = Response,  # Compare by Response
                label = list(
                  Alter ~ "Age",
                  Geschlecht ~ "Sex",
                  Abschluss ~ "Education",
                  T1_BAT_BDI_II_score ~ "BDI-II Score",
                  T1_BAT_BIS_11_score ~ "BIS-11 Score",
                  T1_BAT_CFC_14_score ~ "CFC-14 Score",
                  T1_BAT_FAS_score ~ "FSQ Score",
                  T1_BAT_Kirby_k_score ~ "Kirby k Score",
                  T1_BAT_SRHI_score ~ "SRHI Score",
                  T1_BAT_STAI_T_score ~ "STAI-T Score"
                ),
      statistic = list(
        all_continuous() ~ "{mean} ({sd})",
        all_categorical() ~ "{n} ({p}%)"
      ),
      missing = "no"
    ) %>%
    modify_header(label ~ "**Characteristic**") %>%
    modify_spanning_header(
      starts_with("T1_BAT_") ~ "**Baseline Assessment**"
    ) %>%
    # Remove the row for "Male" in sex
    modify_table_body(
      ~ .x %>% filter(!(variable == "Geschlecht" & label == "Male"))
    )
}

# Loop through the tasks and create tables
characteristics_tables <- lapply(names(task_list), function(task_name) {
  data <- task_list[[task_name]]$subset_data
  generate_characteristics_table(data, task_name)
})

# Name the tables for easy reference
names(characteristics_tables) <- names(task_list)
```

Combined data frame of task-specific descriptive characteristics
```{r}
# Calculate N for Healthy Controls and Patients for each task
calculate_n <- function(data) {
  data %>%
    group_by(Response) %>%
    summarize(N = n(), .groups = "drop") %>%
    pivot_wider(names_from = Response, values_from = N, names_prefix = "Group_") %>%
    rename("Non-Response" = Group_0, "Response" = Group_1)
}

# Get N for each task
stroop_n <- calculate_n(task_list[["Stroop Task"]]$subset_data)
number_letter_n <- calculate_n(task_list[["Number-Letter Task"]]$subset_data)
two_back_n <- calculate_n(task_list[["2-Back Task"]]$subset_data)
ssrt_n <- calculate_n(task_list[["Stop Signal Task"]]$subset_data)

# Convert the tbl_summary tables to tibbles and rename columns dynamically (no information on specific N needed)
stroop_df <- as_tibble(characteristics_tables[["Stroop Task"]]) %>%
  rename_with(~ c("Non-Response", "Response"), .cols = 2:3)

number_letter_df <- as_tibble(characteristics_tables[["Number-Letter Task"]]) %>%
  rename_with(~ c("Non-Response", "Response"), .cols = 2:3)

two_back_df <- as_tibble(characteristics_tables[["2-Back Task"]]) %>%
  rename_with(~ c("Non-Response", "Response"), .cols = 2:3)

ssrt_df <- as_tibble(characteristics_tables[["Stop Signal Task"]]) %>%
  rename_with(~ c("Non-Response", "Response"), .cols = 2:3)

# Insert the "N" row into each data frame
stroop_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Non-Response" = stroop_n[["Non-Response"]], 
         "Response" = stroop_n[["Response"]]),
  stroop_df
)
number_letter_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Non-Response" = number_letter_n[["Non-Response"]], 
         "Response" = number_letter_n[["Response"]]),
  number_letter_df
)
two_back_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Non-Response" = two_back_n[["Non-Response"]], 
         "Response" = two_back_n[["Response"]]),
  two_back_df
)
ssrt_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Non-Response" = ssrt_n[["Non-Response"]], 
         "Response" = ssrt_n[["Response"]]),
  ssrt_df
)

# Combine the modified data frames
characteristics_tasks <- cbind(
  stroop_df[, "**Characteristic**", drop = FALSE],  # Include "Characteristic" column only once
  stroop_df[, -which(names(stroop_df) == "**Characteristic**")],
  number_letter_df[, -which(names(number_letter_df) == "**Characteristic**")],
  two_back_df[, -which(names(two_back_df) == "**Characteristic**")],
  ssrt_df[, -which(names(ssrt_df) == "**Characteristic**")]  # Drop "Characteristic" column from all others
)
```

## Create and save flextable
```{r}
# Rename columns to ensure uniqueness
colnames(characteristics_tasks) <- c(
  "Characteristic", 
  "Stroop_NR", "Stroop_R", 
  "Number_Letter_NR", "Number_Letter_R", 
  "Two_Back_NR", "Two_Back_R", 
  "Stop_Signal_NR", "Stop_Signal_R"
)

# Define primary (task) and secondary (HC vs Pat) headers
primary_headers <- c("Characteristic", "Stroop Task", "Stroop Task", 
                     "Number-Letter Task", "Number-Letter Task", 
                     "2-Back Task", "2-Back Task", 
                     "Stop Signal Task", "Stop Signal Task")

secondary_headers <- c("Characteristic", "Non-Response", "Response", 
                       "Non-Response", "Response", 
                       "Non-Response", "Response", 
                       "Non-Response", "Response")

# Create a flextable
ft <- flextable(characteristics_tasks)

# Add primary and secondary column headers
ft <- set_header_df(
  x = ft,
  mapping = data.frame(
    keys = colnames(characteristics_tasks),
    primary = primary_headers,
    secondary = secondary_headers,
    stringsAsFactors = FALSE
  ),
  key = "keys"
)

# Merge the primary headers to span across related columns
ft <- merge_h(ft, part = "header")

# Align and format the headers
ft <- align(ft, align = "center", part = "header")
ft <- bold(ft, part = "header")
ft <- autofit(ft)

# Write to Word
save_as_docx(ft, path="Characteristics Table Response vs Nonresponse.docx", 
             pr_section = format_flextable_landscape)
```

# Inferential statistics
## Preparation
```{r}
# Ensure "Response" is treated as a factor
data_Pat_pre_clean$Response <- as.factor(data_Pat_pre_clean$Response)

# Define the columns for which to perform the test (i.e., task performance measures)
imp_columns <- c("NumberLetter_BIS_Repeat", "NumberLetter_BIS_Switch", 
                 "NumberLetter_BIS_Diff_Score", "Stroop_BIS_Congruent", 
                 "Stroop_BIS_Incongruent", "Stroop_BIS_Diff_Score", 
                 "TwoBack_BIS_Foil", "TwoBack_BIS_Target", "TwoBack_BIS_Total", 
                 "SSRT")
```

## Test for variance homogeneity
### Define function
```{r}
# Perform Levene's test for homogeneity of variance on task performance measures
levene_test_mult_cols <- function(df_basis, cols) {
  df <- data.frame(p_value = numeric(length(cols)))
  rownames(df) <- cols
  
  for (i in seq_along(cols)) {
    col <- cols[i]
    # Remove NA-cases per variable/task
    df_basis_nomissings <- df_basis[!is.na(df_basis[[col]]),]
    df_basis_nomissings[[col]] <- as.numeric(df_basis_nomissings[[col]])
    # Perform Levene's test, treating 'Response' as a factor
    levene_result <- car::leveneTest(df_basis_nomissings[[col]] ~ df_basis_nomissings[["Response"]])
    df[col, "p_value"] <- round(levene_result[1, "Pr(>F)"], 4)
  }
  
  return(df)
}
```

### Perform levene test
```{r}
levene_test_table <- levene_test_mult_cols(df_basis = data_Pat_pre_clean, cols = imp_columns)
pander(levene_test_table, style = "rmarkdown", fontsize = "tiny")
```
Levene test shows complete variance homoegeneity for all task conditions between responders and non-responders.

## Calculate t-tests 
```{r}
ttest_results_columns <- c(
  "group_mean_Response", "sd_Response",
  "group_mean_NonReponse", "sd_NonResponse",
  "df", "t_statistic", "p_value", "cohen_d", "power"
)

t_test_table <- t_test_mult_cols(df_basis = data_Pat_pre_clean, cols = imp_columns, df_results_columns = ttest_results_columns, grouping_variable = "Response")
#pander(t_test_table, style = "rmarkdown", split.table = Inf, fontsize = "tiny")
t_test_table

# TODO: why is df not an integer?
```
No significant was found for any of the tasks.
Potential reason: low power. Even the largest power for TwoBack_BIS_Foil of 0.22 with a cohen's d of 0.2 would require a sample size of 144 to find a significant effect, as a priori power power analysis shows. However, the patient group is only about half as large as that (n=89), not to speak of the tasks with even smaller power (e.g., NumberLetter_BIS_Repeat with 0.01)

### Prepare table for publication
```{r table preparation, echo=FALSE, warning = FALSE}

t_test_table_pub <- cbind(performance_measures = rownames(t_test_table), t_test_table)

# Combine "Mean" and "SD" columns and reduce dataframe
t_test_table_pub <- t_test_table_pub %>%
  mutate(
    `Responder Mean (SD)` = paste0(round(group_mean_Response, 2), " (", round(sd_Response, 2), ")"),
    `Nonresponder Mean (SD)` = paste0(round(group_mean_NonReponse, 2), " (", round(sd_NonResponse, 2), ")")
    ) %>%
  select(performance_measures, `Responder Mean (SD)`, `Nonresponder Mean (SD)`, df, t_statistic, p_value, p_value_adjusted)

# Rename performance measures using a named vector
rename_map <- c(
  'NumberLetter_BIS_Repeat' = "Number-Letter: Repeat",
  'NumberLetter_BIS_Switch' = "Number-Letter: Switch",
  'NumberLetter_BIS_Diff_Score' = "Number-Letter: Switch - Repeat",
  'Stroop_BIS_Congruent' = "Stroop: Congruent",
  'Stroop_BIS_Incongruent' = "Stroop: Incongruent",
  'Stroop_BIS_Diff_Score' = "Stroop: Incongruent - Congruent",
  'TwoBack_BIS_Foil' = "2-Back: Foil",
  'TwoBack_BIS_Target' = "2-Back: Target",
  'TwoBack_BIS_Total' = "2-Back: Total",
  'SSRT' = "Stop-Signal RT")
t_test_table_pub$performance_measures <- rename_map[t_test_table_pub$performance_measures]

# Change colnames
colnames(t_test_table_pub) <- c("Performance Measure", "Responder Mean (SD)", "Non-Responder Mean (SD)", "df", "t-Statistic", "p-Value", "adj. p-Value")
```

### Create and save flextable
```{r flextable settings, echo=FALSE, warning = FALSE}
ft_t_test <- flextable(t_test_table_pub)

# Set table properties
ft_t_test <- set_table_properties(ft_t_test, width = 1, layout = "autofit")

# Header in bold
ft_t_test <- bold(ft_t_test, bold = TRUE, part = "header")

# Alignments
ft_t_test <- align(ft_t_test, j = 1, align = "left", part = "all") # first column
ft_t_test <- align(ft_t_test, j = 2:ncol(t_test_table_pub), align = "center", part = "all") # rest

# Export flextable
save_as_docx(
  ft_t_test,
  path = "Z:/Projekte_Meinke/Old_projects/Labrotation_Rebecca/Submission/Tables/t-test_Responder_Nonresponder.docx", 
  pr_section = format_flextable_portrait)
```


# Visualization of group differences
## Prepare data
```{r}
# Reshape data from wide to long format, including SSRT, for Response vs. Nonresponse
Response_vs_NonResponse_long <- data_Pat_pre %>%
  pivot_longer(cols = all_of(imp_columns), # imp_columns now includes SSRT
               names_to = "Condition",
               values_to = "BIS_Score") %>%
  drop_na()

# Add a new task variable for faceting
Response_vs_NonResponse_long <- Response_vs_NonResponse_long %>%
  mutate(Task = case_when(
    grepl("NumberLetter", Condition) ~ "Number-Letter Task",
    grepl("Stroop", Condition) ~ "Stroop Task",
    grepl("TwoBack", Condition) ~ "2-Back Task",
    Condition == "SSRT" ~ "Stop Signal Task"
  ))

# Remove prefixes from Condition labels, keeping Task intact
Response_vs_NonResponse_long <- Response_vs_NonResponse_long %>%
  mutate(Condition = gsub("NumberLetter_BIS_|Stroop_BIS_|TwoBack_BIS_", "", Condition)) %>%
  mutate(Condition = as.character(Condition)) %>%
  
  # Set the condition order for each task separately to avoid conflicts (cannot both be "Diff_Score")
  mutate(Condition = ifelse(Task == "NumberLetter" & Condition == "Diff_Score", "Diff_Score NL", Condition)) %>%
  mutate(Condition = ifelse(Task == "Stroop" & Condition == "Diff_Score", "Diff_Score Stroop", Condition)) %>%
  
  # Convert Condition to factor with ordering
  mutate(Condition = case_when(
    Task == "NumberLetter" ~ factor(Condition, levels = c("Repeat", "Switch", "Diff_Score NL")),
    Task == "Stroop" ~ factor(Condition, levels = c("Congruent", "Incongruent", "Diff_Score Stroop")),
    Task == "TwoBack" ~ factor(Condition, levels = c("Target", "Foil", "Total")),
    TRUE ~ as.factor(Condition)
  ))

# Convert "Response" to factor
Response_vs_NonResponse_long$Response <- factor(Response_vs_NonResponse_long$Response, levels = c(0, 1), labels = c("Nonresponders", "Responders"))
```

## Plotting
```{r}
# Create violin plots with mean and rename conditions in the plot

violin_plots_response_vs_nonresponse <- ggplot(Response_vs_NonResponse_long, aes(x = Condition, y = BIS_Score, fill = Response)) +
  geom_violin(position = position_dodge(width = 0.8)) +
  stat_summary(fun.data = mean_sdl, geom = "pointrange", # mult = 2 (SD) by default
               position = position_dodge(width = 0.8)) + # width = 0.65, fatten = 2
  facet_wrap(~ Task, scales = "free") +
  labs(x = "",
       y = "Performance Score",
       fill = "Group") +
  scale_x_discrete(labels = function(x) {
    x[x == "Diff_Score"] <- "Difference \nScore"
    x[x == "Diff_Score Stroop"] <- "Difference \nScore"
    x[x == "Total"] <- "Difference \nScore"
    return(x)
  }) +
  scale_fill_manual(values = c("Nonresponders" = "#D3D3D3", #colors 
                               "Responders" = "#48CFCB")) +
  theme_bw() +
  theme(
    text = element_text(family = "Arial"), #Font
    axis.title = element_text(size = 24),
    axis.title.y = element_text(size = 24, margin = margin(r = 30)),
    axis.text = element_text(size = 18),
    legend.title = element_text(size = 24),
    legend.text = element_text(size = 18),
    strip.text = element_text(size = 20),
    panel.spacing.x = unit(3, "lines")
  )
print(violin_plots_response_vs_nonresponse)

ggsave(paste0("response_vs_nonresponse",param_suffix,".svg"),violin_plots_response_vs_nonresponse,
       width = 10)
```
