---
title: "Group Comparison: Healthy Controls vs. Patients"
authors: "Charlotte Meinke, Rebecca Delfendahl, Till Julius Adam"
date: "2024-09-04"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
params:
  input_data_path: "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/0_Datapreparation/RT_trimmed_RT_wrong_removed/BIS/outliers-removed"
---
<style type="text/css">
.main-container { /* Adjust main blocks */
  max-width: 100% !important;
  margin: auto;
}

body {
  font-family: "Georgia", serif !important; /* Set the font for the entire document */
}

.tocify { /* Adjust table of contents */
  max-width: 100% !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(pander)
library(effsize)
library(pwr)
library(gtsummary)
library(gt)
library(car)
library(flextable)
library(officer)
library(broom)
library(knitr)

# User-defined functions
source("../Useful_functions.R")
```

# Set-Up

### 1. Settings flextable
```{r}
format_flextable_portrait <- flextable_settings(word_orientation = "portrait")
format_flextable_landscape <- flextable_settings(word_orientation = "landscape")
```

### 2. Importing Data
```{r}
basic_path = params$input_data_path
data_HC <- read.csv(file.path(basic_path,"Data_HC.csv"))
data_Pat_pre <- read.csv(file.path(basic_path,"Data_Patients_Pre.csv"))
```

### 3. Set results path to save tables and images
```{r}
results_path <- create_results_path(inputdata_path = basic_path, 
                                      output_mainpath = "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/1_Group_comparison/HC_vs_Pat"
                                          )
```

### 4. Prepare dataset
```{r}
## 1. Merge patients pre and controls
Patients_vs_HC <- rbind(data_Pat_pre, data_HC)

## 2. Reduce dataset to executive functions of interest (BIS columns and SSRT)
BIS_columns <- colnames(Patients_vs_HC)[grep("BIS", colnames(Patients_vs_HC))]
BIS_columns <- BIS_columns[BIS_columns != "T1_BAT_BIS_11_score"] # BIS-11 is a questionnaire
imp_columns <- c(BIS_columns,"SSRT")
Patients_vs_HC_imp <- Patients_vs_HC[,c("Subject","Gruppe",imp_columns)]

## 3. Add confounds (Alter, Geschlecht, Abschluss)
Patients_vs_HC_imp_conf <- merge(Patients_vs_HC[,c("Subject","Alter","Geschlecht","Abschluss")], Patients_vs_HC_imp, by = "Subject")
```
In total, data for `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1,])` patients and `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0,])` healthy controls were available. When outliers were removed, the number of subjects varied slightly: 
Number-Letter Task: `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0 & !is.na(Patients_vs_HC_imp$NumberLetter_BIS_Repeat),])` healthy controls,  `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1 & !is.na(Patients_vs_HC_imp$NumberLetter_BIS_Repeat),])` patients;  
Stroop Task: `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0 & !is.na(Patients_vs_HC_imp$Stroop_BIS_Congruent),])` healthy controls,  `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1 & !is.na(Patients_vs_HC_imp$Stroop_BIS_Congruent),])` patients;  
2back-Task: `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0 & !is.na(Patients_vs_HC_imp$TwoBack_BIS_Total),])` healthy controls,  `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1 & !is.na(Patients_vs_HC_imp$TwoBack_BIS_Total),])` patients;  

SSRT: `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0 & !is.na(Patients_vs_HC_imp$SSRT),])` healthy controls,  `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1 & !is.na(Patients_vs_HC_imp$SSRT),])` patients.


# Inferential statistics

## Main analysis

### 1. Test for variance homogeneity
```{r}
# Ensure "Gruppe" is treated as a factor
Patients_vs_HC_imp$Gruppe <- as.factor(Patients_vs_HC_imp$Gruppe)

# Perform Levene's test for homogeneity of variance
levene_test_table <- levene_test_mult_cols(df_basis = Patients_vs_HC_imp, 
                                           cols = imp_columns, 
                                           grouping_variable = "Gruppe")
#pander(levene_test_table, style = "rmarkdown", fontsize = "tiny")
levene_test_table
```

### 2. Calculate Welch t-test
```{r}
t_test_table <- t_test_mult_cols(df_basis = Patients_vs_HC_imp, cols = imp_columns, grouping_variable = "Gruppe")
#pander(t_test_table, style = "rmarkdown", fontsize = "tiny")
t_test_table
```

### 3. Prepare table for publication
```{r}
prepare_ttest_table <- function(ttest_table, dv_name){
  # Convert rownames (dependent variables) to separate column
  t_test_table_pub <- cbind(dependent_variables = rownames(ttest_table), ttest_table)
  
  # Combine "Mean" and "SD" columns and reduce dataframe
  t_test_table_pub <- t_test_table_pub %>%
  mutate(
    `Patients Mean (SD)` = paste0(round(group_1_mean, 2), " (", round(group_1_sd, 2), ")"),
    `HC Mean (SD)` = paste0(round(group_0_mean, 2), " (", round(group_0_mean, 2), ")")
    ) %>%
  select(dependent_variables, `Patients Mean (SD)`, `HC Mean (SD)`, 
         df, t_statistic, effect_size, p_value, p_value_adjusted)
  
  # Change colnames
  colnames(t_test_table_pub) <- c(dv_name, "Patients Mean (SD)", "HC Mean (SD)", "df", "t-Statistic", "Hedges` g", "p-Value", "adj. p-Value")
  
  return(t_test_table_pub)
}

t_test_table_pub <- prepare_ttest_table(ttest_table = t_test_table, dv_name = "Performance Measure")

# Rename performance measures using a named vector
rename_map <- c(
  'NumberLetter_BIS_Repeat' = "Number-Letter: Repeat",
  'NumberLetter_BIS_Switch' = "Number-Letter: Switch",
  'NumberLetter_BIS_Diff_Score' = "Number-Letter: Switch - Repeat",
  'Stroop_BIS_Congruent' = "Stroop: Congruent",
  'Stroop_BIS_Incongruent' = "Stroop: Incongruent",
  'Stroop_BIS_Diff_Score' = "Stroop: Incongruent - Congruent",
  'TwoBack_BIS_Foil' = "2-Back: Foil",
  'TwoBack_BIS_Target' = "2-Back: Target",
  'TwoBack_BIS_Total' = "2-Back: Total",
  'SSRT' = "Stop-Signal RT")
t_test_table_pub$dependent_variables <- rename_map[t_test_table_pub$dependent_variables]
```

### 4. Create and save flextable
```{r}
ft_t_test <- flextable(t_test_table_pub)

# Set table properties
ft_t_test <- set_table_properties(ft_t_test, width = 1, layout = "autofit")

# Header in bold
ft_t_test <- bold(ft_t_test, bold = TRUE, part = "header")

# Alignments
ft_t_test <- align(ft_t_test, j = 1, align = "left", part = "all") # first column
ft_t_test <- align(ft_t_test, j = 2:ncol(t_test_table_pub), align = "center", part = "all") # rest

# Export flextable
save_as_docx(
  ft_t_test,
  path = file.path(results_path, "t-test_Patients_HCs.docx"), 
  pr_section = format_flextable_portrait)
```

## Descriptive group differences

### 0. Identify categorical vs. dimensional variables
```{r}
categorical_vars <- c("Abschluss", "Geschlecht")
dimensional_vars <- c("Alter", "T1_BAT_FAS_score", "T1_BAT_BDI_II_score", "T1_BAT_STAI_T_score", "T1_BAT_BIS_11_score", "T1_BAT_Kirby_k_score", "T1_BAT_CFC_14_score", "T1_BAT_SRHI_score")
```

### 1. Welch t-test for dimensional variables
```{r}
t_test_descr <- t_test_mult_cols(df_basis = Patients_vs_HC, cols = dimensional_vars, grouping_variable = "Gruppe")

t_test_descr_pub <- prepare_ttest_table(ttest_table = t_test_descr, dv_name = "Dimensional Variables")
```

### 2. Chi-square test for categorical variables
```{r}
chi_sq_test_descr <- chi_sq_test_mult_cols(df_basis = Patients_vs_HC, cols = categorical_vars, grouping_variable = "Gruppe")
```


# Plotting

### 1. Prepare data
```{r}
# Reshape data from wide to long format, including SSRT
Patients_vs_HC_long <- Patients_vs_HC_imp %>%
  pivot_longer(cols = all_of(imp_columns), # imp_columns now includes SSRT
               names_to = "Condition",
               values_to = "BIS_Score")

# Add new task variable to keep the headers intact for faceting
Patients_vs_HC_long <- Patients_vs_HC_long %>%
  mutate(Task = case_when(
    grepl("NumberLetter", Condition) ~ "Number-Letter Task",
    grepl("Stroop", Condition) ~ "Stroop Task",
    grepl("TwoBack", Condition) ~ "2-Back Task",
    Condition == "SSRT" ~ "Stop Signal Task"
  ))

# Remove prefixes from Condition labels, keeping Task intact
Patients_vs_HC_long <- Patients_vs_HC_long %>%
  mutate(Condition = gsub("NumberLetter_BIS_|Stroop_BIS_|TwoBack_BIS_", "", Condition)) %>%
  mutate(Condition = as.character(Condition)) %>%
  
  # Set the condition order for each task separately to avoid conflicts (cannot both be "Diff_Score")
  mutate(Condition = ifelse(Task == "NumberLetter" & Condition == "Diff_Score", "Diff_Score NL", Condition)) %>%
  mutate(Condition = ifelse(Task == "Stroop" & Condition == "Diff_Score", "Diff_Score Stroop", Condition)) %>%
  
  # Convert Condition to factor with ordering
  mutate(Condition = case_when(
    Task == "NumberLetter" ~ factor(Condition, levels = c("Repeat", "Switch", "Diff_Score NL")),
    Task == "Stroop" ~ factor(Condition, levels = c("Congruent", "Incongruent", "Diff_Score Stroop")),
    Task == "TwoBack" ~ factor(Condition, levels = c("Foil", "Target", "Total")),
    TRUE ~ as.factor(Condition)
  ))

# Convert "Gruppe" to factor
Patients_vs_HC_long$Gruppe <- factor(Patients_vs_HC_long$Gruppe, levels = c(0, 1), labels = c("Healthy Controls", "Patients"))
```

### 2. Violin Plot: Patients vs. Healthy Controls
```{r, fig.width = 16, fig.height = 9, warning = FALSE}
# Create violin plots with mean and rename conditions in the plot

violin_plots_patients_hc <- ggplot(Patients_vs_HC_long, aes(x = Condition, y = BIS_Score, fill = Gruppe)) +
  geom_violin(position = position_dodge(width = 0.8)) +
  stat_summary(fun.data = mean_sdl, geom = "pointrange", # mult = 2 (SD) by default
               position = position_dodge(width = 0.8)) + # width = 0.65, fatten = 2
  facet_wrap(~ Task, scales = "free") +
  labs(x = "",
       y = "Performance Score",
       fill = "Group") +
  scale_x_discrete(labels = function(x) {
    x[x == "Diff_Score"] <- "Difference \nScore"
    x[x == "Diff_Score Stroop"] <- "Difference \nScore"
    return(x)
  }) +
  scale_fill_manual(values = c("Healthy Controls" = "#D3D3D3", #colors 
                                 "Patients" = "#48CFCB")) +
  theme_bw() +
  theme(
    text = element_text(family = "Arial"), #Font
    axis.title = element_text(size = 24),
    axis.title.y = element_text(size = 24, margin = margin(r = 30)),
    axis.text = element_text(size = 18),
    legend.title = element_text(size = 24),
    legend.text = element_text(size = 18),
    strip.text = element_text(size = 20),
    panel.spacing.x = unit(3, "lines")
  )
print(violin_plots_patients_hc)
ggsave(file.path(results_path, "violin_plots_patients_hc.png"), 
       plot = violin_plots_patients_hc, width = 16, height = 9, dpi = 300)
ggsave(file.path(results_path, "violin_plots_patients_hc.svg"), 
       plot = violin_plots_patients_hc, width = 16, height = 9, dpi = 300)
```

### 3. Boxplot: Patients vs. Healthy Controls
```{r, warning = FALSE}
# Create boxplots
boxplots_Patients_HC <- ggplot(Patients_vs_HC_long, aes(x = Condition, y = BIS_Score, color = Gruppe)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(position = position_jitterdodge(),
              size = 0.7,
              shape = 21,
              fill = NA) +
  facet_wrap(~ Task, scales = "free") +
  #facet_grid(rows = vars(Task), scales = "free_x") +
  labs(title = "Boxplots of Task Condition by Group",
       x = "Condition",
       y = "BIS-Score",
       color = "Group")
print(boxplots_Patients_HC)
ggsave(file.path(results_path, "boxplots_patients_hc.png"), 
       plot = boxplots_Patients_HC, width = 16, height = 9, dpi = 300)
```
