---
title: "Group Comparison: Pre Post"
authors: "Charlotte Meinke, Rebecca Delfendahl, Till Julius Adam"
date: "2024-11-20"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
params:
  input_data_path: "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/0_Datapreparation/RT_trimmed_RT_wrong_removed/BIS/outliers-removed"
---
<style type="text/css">
.main-container { /* Adjust main blocks */
  max-width: 100% !important;
  margin: auto;
}

body {
  font-family: "Georgia", serif !important; /* Set the font for the entire document */
}

.tocify { /* Adjust table of contents */
  max-width: 100% !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(pander)
library(effsize)
library(pwr)
library(stats)
library(gtsummary)
library(gt)
library(car)
library(flextable)
library(officer)
library(broom)
library(knitr)

# User-defined functions
source("../Useful_functions.R")
```

## <u>Importing Data</u>

```{r}
basic_path = params$input_data_path
data_HC <- read.csv(file.path(basic_path,"Data_HC.csv"))
data_Pat_pre <- read.csv(file.path(basic_path,"Data_Patients_Pre.csv"))
data_Pat_post <- read.csv(file.path(basic_path,"Data_Patients_Post.csv"))
```

### Prepare dataset
```{r}
## 1. Merge patients pre and controls
Patients_vs_HC <- rbind(data_Pat_pre, data_HC)

## 2. Reduce dataset to executive functions of interest (BIS columns and SSRT)
BIS_columns <- colnames(Patients_vs_HC)[grep("BIS", colnames(Patients_vs_HC))]
BIS_columns <- BIS_columns[BIS_columns != "T1_BAT_BIS_11_score"] # BIS-11 is a questionnaire
imp_columns <- c(BIS_columns, "SSRT")
Patients_vs_HC_imp <- Patients_vs_HC[,c("Subject","Gruppe",imp_columns)]

# 3. Identify common subjects in both datasets
data_Pat_pre$Subject_postID <- data_Pat_pre$Subject + 600 # IDs at post = ID at pre + 600

common_subjects_postID <- intersect(data_Pat_pre$Subject_postID, data_Pat_post$Subject) #Using intersect function to exclude only pre/post cases

# 4. Filter the datasets to only include common subjects
data_Pat_pre_filtered <- data_Pat_pre[data_Pat_pre$Subject_postID %in% common_subjects_postID , ]
data_Pat_post_filtered <- data_Pat_post[data_Pat_post$Subject %in% common_subjects_postID , ]

# 5. Important columns
data_Pat_pre_filtered_imp <- data_Pat_pre_filtered[, c("Subject", imp_columns)]
data_Pat_post_filtered_imp <- data_Pat_post_filtered[, c("Subject", imp_columns)]
```
Pre and post measures were available for `r nrow(data_Pat_pre_filtered_imp)` patients.

### Testing Assumption of dependent sample t-test: Positive correlation of measurement series
```{r}
## Calculate Raw Pearson Correlation and Test Significance
calculate_correlations_and_significance <- function(pre_data, post_data, cols) {
  df <- data.frame(Correlation = numeric(length(cols)), t_statistic = numeric(length(cols)), 
                   p_value = numeric(length(cols)))
  rownames(df) <- cols
  p_values_raw <- numeric(length(cols))
  for (i in seq_along(cols)) {
    col <- cols[i]
    
    # Use complete.cases to keep only rows where both pre AND post are not NA
    complete_cases <- complete.cases(pre_data[[col]], post_data[[col]])
    pre_col_data <- pre_data[[col]][complete_cases]
    post_col_data <- post_data[[col]][complete_cases]
    
    # Calculate Pearson correlation
    correlation <- cor(pre_col_data, post_col_data)
    
    # Calculate the t-statistic for the correlation
    n <- length(pre_col_data)
    t_statistic <- correlation * sqrt((n - 2) / (1 - correlation^2))
    
    # Calculate the p-value for a one-tailed test (H_0: r ≤ 0)
    p_value <- pt(t_statistic, df = n - 2, lower.tail = FALSE)
    p_values_raw[i] <- p_value
    
    # Store the results
    df[col, "Correlation"] <- round(correlation, 3)
    df[col, "t_statistic"] <- round(t_statistic, 3)
    df[col, "p_value"] <- round(p_value, 3)
  }
  
  # Adjust p-values using Benjamini-Hochberg method
  p_values_adjusted <- p.adjust(p_values_raw, method = "BH")
  df$p_value_adjusted <- round(p_values_adjusted, 2)
  
  return(df)
}


# Apply the correlation and significance test
correlation_significance_table <- calculate_correlations_and_significance(pre_data = data_Pat_pre_filtered_imp, post_data = data_Pat_post_filtered_imp, cols = imp_columns)
pander(correlation_significance_table, style = "rmarkdown", fontsize = "tiny")
```
For all tasks, the correlations of the measurement series are significantly positive. Thus, the assumption is met for all tasks. 

### Calculate Welch t-test
```{r}
paired_t_test_mult_cols <- function(pre_data, post_data, cols) {
  df <- data.frame(t_statistic = numeric(length(cols)), p_value = numeric(length(cols)),
                   group_mean_pre = numeric(length(cols)), group_mean_post = numeric(length(cols)),
                   cohen_d = numeric(length(cols)))
  rownames(df) <- cols
  p_values_raw <- numeric(length(cols)) # Vector to store raw p-values
  
  for (i in seq_along(cols)) {
    col <- cols[i]
    
    # Use complete.cases to keep only rows where both pre AND post are not NA
    complete_cases <- complete.cases(pre_data[[col]], post_data[[col]])
    pre_col_data <- pre_data[[col]][complete_cases]
    post_col_data <- post_data[[col]][complete_cases]
    
    # Check if there are at least 2 complete pairs of observations
    if (length(pre_col_data) < 2 || length(post_col_data) < 2) {
      warning(paste("Not enough data for column:", col))
      df[col, ] <- NA  # Assign NA to the row in case of insufficient data
      next  # Skip to the next column
    }
    
    # Perform paired Welch t-test (non-formula interface)
    results <- t.test(pre_col_data, post_col_data, paired = TRUE, var.equal = FALSE)
    #Alternativ using formula method:
    #results <- t.test(df_basis[[col]] ~ df_basis[["time"]], paired = TRUE, var.equal = FALSE)
    
    # Store raw p-values for BH correction
    p_values_raw[i] <- results$p.value
    
    # Calculate Cohen's d for paired samples (effect size)
    cohen_d_result <- cohen.d(post_col_data, pre_col_data, paired = TRUE)
    cohen_d <- cohen_d_result$estimate
    
    # Calculate power using pwr.t.test
    n <- length(pre_col_data)
    power_result <- pwr.t.test(d = cohen_d, n = n, sig.level = 0.05, type = "paired", alternative = "greater")
    power <- power_result$power
    
    # Store the results
    df[col, "t_statistic"] <- round(results$statistic, 2)
    df[col, "p_value"] <- round(results$p.value, 2)
    df[col, "group_mean_pre"] <- round(mean(pre_col_data), 2)
    df[col, "sd_pre"] <- round(sd(pre_col_data), 2)
    df[col, "group_mean_post"] <- round(mean(post_col_data), 2)
    df[col, "sd_post"] <- round(sd(post_col_data), 2)
    df[col, "cohen_d"] <- round(cohen_d, 2)
    df[col, "power"] <- round(power, 2)
  }
  
  # Adjust p-values using Benjamini-Hochberg method for multiple testing of related tasks
  p_values_adjusted <- p.adjust(p_values_raw, method = "BH")
  df$p_value_adjusted <- round(p_values_adjusted, 2)
  
  return(df)
}

# Apply the paired Welch t-test to your data
t_test_table_pre_post <- paired_t_test_mult_cols(pre_data = data_Pat_pre_filtered_imp, post_data = data_Pat_post_filtered_imp, cols = imp_columns)
pander(t_test_table_pre_post)
```
Patients performed better in all measurements of executive functioning after the exposure session. This might be due to a practice effect.

(Manual check)
```{r}
# Number of patients in the pre group
n_pre <- nrow(data_Pat_pre)

# Number of patients in the post group
n_post <- nrow(data_Pat_post)

# Output the results
n_pre
n_post
```

## Boxplot
```{r}
# Add a timepoint variable to the filtered datasets and combine
data_Pat_pre_filtered_imp$time <- "pre"
data_Pat_post_filtered_imp$time <- "post"
Patients_pre_vs_post_long <- rbind(data_Pat_pre_filtered_imp, data_Pat_post_filtered_imp)

# Reshape data from wide to long format
Patients_pre_vs_post_long <- Patients_pre_vs_post_long %>%
  pivot_longer(cols = all_of(BIS_columns), # imp_columns to include SSRT
               names_to = "Condition",
               values_to = "BIS_Score")

# Add new task variable
Patients_pre_vs_post_long <- Patients_pre_vs_post_long %>%
  mutate(Task = case_when(
    grepl("NumberLetter", Condition) ~ "NumberLetter",
    grepl("Stroop", Condition) ~ "Stroop",
    grepl("TwoBack", Condition) ~ "TwoBack"
  ))

# Remove prefixes from Condition labels, keeping Task intact
Patients_pre_vs_post_long <- Patients_pre_vs_post_long %>%
  mutate(Condition = gsub("NumberLetter_BIS_|Stroop_BIS_|TwoBack_BIS_", "", Condition))

# Convert "time" to factor, specify order
Patients_pre_vs_post_long$time <- factor(Patients_pre_vs_post_long$time, levels = c("pre", "post"))

# Create boxplot with paired observations
boxplot_Patients_pre_post <- ggplot(Patients_pre_vs_post_long, aes(x = Condition, y = BIS_Score, color = time)) + # group = Subject
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(position = position_jitterdodge(),
              size = 0.7,
              shape = 21,
              fill = NA) +
  #geom_line(aes(group = Subject)) + # optional, if you want to visualize pairing
  facet_wrap(~ Task, scales = "free_x") +
  labs(title = "Boxplots of Task Condition by Time of Measurement",
       x = "Condition",
       y = "BIS-Score",
       color = "Time")

print(boxplot_Patients_pre_post)
```