---
title: "Group Comparison: Healthy Controls vs. Patients"
authors: "Charlotte Meinke, Rebecca Delfendahl, Till Julius Adam"
date: "2024-09-04"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
params:
  input_data_path: "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/0_Datapreparation/RT_trimmed_RT_wrong_removed/BIS/outliers-removed"
---
<style type="text/css">
.main-container { /* Adjust main blocks */
  max-width: 100% !important;
  margin: auto;
}

body {
  font-family: "Georgia", serif !important; /* Set the font for the entire document */
}

.tocify { /* Adjust table of contents */
  max-width: 100% !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(pander)
library(effsize)
library(pwr)
library(gtsummary)
library(gt)
library(car)
library(flextable)
library(officer)
library(broom)
library(knitr)

# User-defined functions
source("../Useful_functions.R")
```

# Set-Up
### 1. Settings flextable
```{r}
format_flextable_portrait <- flextable_settings(word_orientation = "portrait")
format_flextable_landscape <- flextable_settings(word_orientation = "landscape")
```

### 2. Importing Data
```{r}
basic_path = params$input_data_path
data_HC <- read.csv(file.path(basic_path,"Data_HC.csv"))
data_Pat_pre <- read.csv(file.path(basic_path,"Data_Patients_Pre.csv"))
```

### 3. Set results path to save tables and images
```{r}
results_path <- create_results_path(inputdata_path = basic_path, 
                                      output_mainpath = "Y:/PsyThera/Projekte_Meinke/Old_projects/Labrotation_Rebecca/1_Group_comparison/HC_vs_Pat"
                                          )
```

### 4. Prepare dataset
```{r}
## 1. Merge patients pre and controls
Patients_vs_HC <- rbind(data_Pat_pre, data_HC)

## 2. Reduce dataset to executive functions of interest (BIS columns and SSRT)
BIS_columns <- colnames(Patients_vs_HC)[grep("BIS", colnames(Patients_vs_HC))]
BIS_columns <- BIS_columns[BIS_columns != "T1_BAT_BIS_11_score"] # BIS-11 is a questionnaire
imp_columns <- c(BIS_columns,"SSRT")
Patients_vs_HC_imp <- Patients_vs_HC[,c("Subject","Gruppe",imp_columns)]

## 3. Add confounds (Alter, Geschlecht, Abschluss)
Patients_vs_HC_imp_conf <- merge(Patients_vs_HC[,c("Subject","Alter","Geschlecht","Abschluss")], Patients_vs_HC_imp, by = "Subject")
```
In total, data for `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1,])` patients and `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0,])` healthy controls were available. When outliers were removed, the number of subjects varied slightly: 
Number-Letter Task: `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0 & !is.na(Patients_vs_HC_imp$NumberLetter_BIS_Repeat),])` healthy controls,  `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1 & !is.na(Patients_vs_HC_imp$NumberLetter_BIS_Repeat),])` patients;  
Stroop Task: `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0 & !is.na(Patients_vs_HC_imp$Stroop_BIS_Congruent),])` healthy controls,  `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1 & !is.na(Patients_vs_HC_imp$Stroop_BIS_Congruent),])` patients;  
2back-Task: `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0 & !is.na(Patients_vs_HC_imp$TwoBack_BIS_Total),])` healthy controls,  `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1 & !is.na(Patients_vs_HC_imp$TwoBack_BIS_Total),])` patients;  

SSRT: `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 0 & !is.na(Patients_vs_HC_imp$SSRT),])` healthy controls,  `r nrow(Patients_vs_HC_imp[Patients_vs_HC_imp$Gruppe == 1 & !is.na(Patients_vs_HC_imp$SSRT),])` patients.

# Descriptive Statistics
### 1. Prepare task-specific data frame of characteristics
```{r}
# 1. Subset data frame by non-NA for respective tasks
task_list <- list(
  "Stroop Task" = list(
    subset_data = Patients_vs_HC %>%
      filter(!is.na(Stroop_BIS_Congruent) & 
             !is.na(Stroop_BIS_Incongruent) & 
             !is.na(Stroop_BIS_Diff_Score))
  ),
  "Number-Letter Task" = list(
    subset_data = Patients_vs_HC %>%
      filter(!is.na(NumberLetter_BIS_Repeat) & 
             !is.na(NumberLetter_BIS_Switch) & 
             !is.na(NumberLetter_BIS_Diff_Score))
  ),
  "2-Back Task" = list(
    subset_data = Patients_vs_HC %>%
      filter(!is.na(TwoBack_BIS_Foil) & 
             !is.na(TwoBack_BIS_Target) & 
             !is.na(TwoBack_BIS_Total))
  ),
  "Stop Signal Task" = list(
    subset_data = Patients_vs_HC %>%
      filter(!is.na(SSRT))
  )
)


# 2. Function to generate characteristics table
generate_characteristics_table <- function(data, task_name) {
  data %>%
    select(Gruppe, Alter, Geschlecht, Abschluss, 
           T1_BAT_BDI_II_score, T1_BAT_BIS_11_score, T1_BAT_CFC_14_score, 
           T1_BAT_FAS_score, T1_BAT_Kirby_k_score, T1_BAT_SRHI_score, T1_BAT_STAI_T_score) %>%
    mutate(Gruppe = factor(Gruppe, levels = c(0, 1), labels = c("Healthy Controls", "Patients")),
           Geschlecht = factor(Geschlecht, levels = c(0, 1), labels = c("Male", "Female")),
           Abschluss = factor(Abschluss, levels = c(1, 2, 3, 4), 
                              labels = c("Basic secondary education", 
                                         "Intermediate secondary education",
                                         "Higher secondary education", 
                                         "Other"))) %>%
    tbl_summary(by = Gruppe,  # Compare by Gruppe (HC vs. Pat),
                digits = list(T1_BAT_BDI_II_score = c(1, 1),
                              T1_BAT_BIS_11_score = c(1, 1),
                              T1_BAT_CFC_14_score = c(1, 1),
                              T1_BAT_FAS_score = c(1, 1),
                              T1_BAT_SRHI_score = c(1, 1),
                              T1_BAT_STAI_T_score= c(1, 1)),
                label = list(
                  Alter ~ "Age",
                  Geschlecht ~ "Sex",
                  Abschluss ~ "Education",
                  T1_BAT_BDI_II_score ~ "BDI-II Score",
                  T1_BAT_BIS_11_score ~ "BIS-11 Score",
                  T1_BAT_CFC_14_score ~ "CFC-14 Score",
                  T1_BAT_FAS_score ~ "FSQ Score",
                  T1_BAT_Kirby_k_score ~ "Kirby k Score",
                  T1_BAT_SRHI_score ~ "SRHI Score",
                  T1_BAT_STAI_T_score ~ "STAI-T Score"
                ),
      statistic = list(
        all_continuous() ~ "{mean} ({sd})",
        all_categorical() ~ "{n} ({p}%)"
      ),
      missing = "ifany"
    ) %>%
    modify_header(label ~ "**Characteristic**") %>%
    modify_spanning_header(
      starts_with("T1_BAT_") ~ "**Baseline Assessment**"
    ) %>%
    # Remove the row for "Male" in sex
    modify_table_body(
      ~ .x %>% filter(!(variable == "Geschlecht" & label == "Male"))
    ) 
}

# Loop through the tasks and create tables
characteristics_tables <- lapply(names(task_list), function(task_name) {
  data <- task_list[[task_name]]$subset_data
  generate_characteristics_table(data, task_name)
})

# Name the tables for easy reference
names(characteristics_tables) <- names(task_list)


# 3. Combined data frame of task-specific descriptive characteristics
# Calculate N for Healthy Controls and Patients for each task
calculate_n <- function(data) {
  data %>%
    group_by(Gruppe) %>%
    summarize(N = n(), .groups = "drop") %>%
    pivot_wider(names_from = Gruppe, values_from = N, names_prefix = "Group_") %>%
    rename("Healthy Controls" = Group_0, "Patients" = Group_1)
}

# Get N for each task
stroop_n <- calculate_n(task_list[["Stroop Task"]]$subset_data)
number_letter_n <- calculate_n(task_list[["Number-Letter Task"]]$subset_data)
two_back_n <- calculate_n(task_list[["2-Back Task"]]$subset_data)
ssrt_n <- calculate_n(task_list[["Stop Signal Task"]]$subset_data)

# Convert the tbl_summary tables to tibbles and rename columns dynamically (no information on specific N needed)
stroop_df <- as_tibble(characteristics_tables[["Stroop Task"]]) %>%
  rename_with(~ c("Healthy Controls", "Patients"), .cols = 2:3)

number_letter_df <- as_tibble(characteristics_tables[["Number-Letter Task"]]) %>%
  rename_with(~ c("Healthy Controls", "Patients"), .cols = 2:3)

two_back_df <- as_tibble(characteristics_tables[["2-Back Task"]]) %>%
  rename_with(~ c("Healthy Controls", "Patients"), .cols = 2:3)

ssrt_df <- as_tibble(characteristics_tables[["Stop Signal Task"]]) %>%
  rename_with(~ c("Healthy Controls", "Patients"), .cols = 2:3)

# Insert the "N" row into each data frame
stroop_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Healthy Controls" = stroop_n[["Healthy Controls"]], 
         "Patients" = stroop_n[["Patients"]]),
  stroop_df
)
number_letter_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Healthy Controls" = number_letter_n[["Healthy Controls"]], 
         "Patients" = number_letter_n[["Patients"]]),
  number_letter_df
)
two_back_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Healthy Controls" = two_back_n[["Healthy Controls"]], 
         "Patients" = two_back_n[["Patients"]]),
  two_back_df
)
ssrt_df <- rbind(
  tibble("**Characteristic**" = "N", 
         "Healthy Controls" = ssrt_n[["Healthy Controls"]], 
         "Patients" = ssrt_n[["Patients"]]),
  ssrt_df
)

# Combine the modified data frames
characteristics_tasks <- cbind(
  stroop_df[, "**Characteristic**", drop = FALSE],  # Include "Characteristic" column only once
  stroop_df[, -which(names(stroop_df) == "**Characteristic**")],
  number_letter_df[, -which(names(number_letter_df) == "**Characteristic**")],
  two_back_df[, -which(names(two_back_df) == "**Characteristic**")],
  ssrt_df[, -which(names(ssrt_df) == "**Characteristic**")]  # Drop "Characteristic" column from all others
)
```

### 2. Flextable of characteristics by task
```{r}
# Create Flextable
# Rename columns to ensure uniqueness
colnames(characteristics_tasks) <- c(
  "Characteristic", 
  "Stroop_Healthy_Controls", "Stroop_Patients", 
  "Number_Letter_Healthy_Controls", "Number_Letter_Patients", 
  "Two_Back_Healthy_Controls", "Two_Back_Patients", 
  "Stop_Signal_Healthy_Controls", "Stop_Signal_Patients"
)

# Define primary (task) and secondary (HC vs Pat) headers
primary_headers <- c("Characteristic", "Stroop Task", "Stroop Task", 
                     "Number-Letter Task", "Number-Letter Task", 
                     "2-Back Task", "2-Back Task", 
                     "Stop Signal Task", "Stop Signal Task")

secondary_headers <- c("Characteristic", "Healthy Controls", "Patients", 
                       "Healthy Controls", "Patients", 
                       "Healthy Controls", "Patients", 
                       "Healthy Controls", "Patients")

# Create a flextable
ft <- flextable(characteristics_tasks)

# Add primary and secondary column headers
ft <- set_header_df(
  x = ft,
  mapping = data.frame(
    keys = colnames(characteristics_tasks),
    primary = primary_headers,
    secondary = secondary_headers,
    stringsAsFactors = FALSE
  ),
  key = "keys"
)

# Merge the primary headers to span across related columns
ft <- merge_h(ft, part = "header")

# Align and format the headers
ft <- align(ft, align = "center", part = "header")
ft <- bold(ft, part = "header")
ft <- autofit(ft)

# Write to Word
save_as_docx(
  ft,
  path =file.path(results_path, "Characteristics_Patients_HCs.docx"), 
  pr_section = format_flextable_portrait)
ft
```

## 3. Get information about removed trials during the calculation of the mean RT
```{r}
library(dplyr)

# Select columns
columns_n_trimmed <- grep("_n_trimmed_removed_RT$", names(Patients_vs_HC), value = TRUE)
columns_n_wrong <- grep("_n_wrong_removed_RT$", names(Patients_vs_HC), value = TRUE)

# Create a data frame dynamically
df <- data.frame(
  Condition = gsub("_n_trimmed_removed_RT", "", columns_n_trimmed),  # Remove suffix for clean names
  n_trimmed_removed = sapply(Patients_vs_HC[ columns_n_trimmed], mean, na.rm = TRUE),
  n_trimmed_removed_sd = sapply(Patients_vs_HC[ columns_n_trimmed], sd, na.rm = TRUE),
  n_wrong_removed = sapply(Patients_vs_HC[columns_n_wrong], mean, na.rm = TRUE),
  n_wrong_removed_sd = sapply(Patients_vs_HC[columns_n_wrong], sd, na.rm = TRUE)
)

# Rename variables
df$Condition <- dplyr::recode(df$Condition, Repeat = "Number-Letter: Repeat",
                          Switch = "Number-Letter: Switch", Congruent = "Stroop: Congruent",
                          Incongruent = "Stroop: Incongruent", Foil = "2-Back: Foil", 
                          Target = "2-Back: Target")
colnames(df) <- c("Condition", "Number of trials removed due to trimming (Mean)",
                  "Number of trials removed due to trimming (SD)",
                  "Number of wrong trials removed (Mean)",
                  "Number of wrong trials removed (SD)")

# Turn to flextable
ft_removedRT <- flextable(df)

# Round to 2
ft_removedRT <- colformat_double(ft_removedRT, digits = 2)

# Set table properties
ft_removedRT <- set_table_properties(ft_removedRT, width = 1, layout = "autofit")

# Header in bold
ft_removedRT <- bold(ft_removedRT, bold = TRUE, part = "header")

# Alignments
ft_removedRT<- align(ft_removedRT, j = 1, align = "left", part = "all") # first column
ft_removedRT <- align(ft_removedRT, j = 2:ncol(df), align = "center", part = "all") # rest

# Export flextable
save_as_docx(
  ft_removedRT,
  path = file.path(results_path, "removed_trials_Patients_HCs.docx"), 
  pr_section = format_flextable_portrait)

# Info for interpreting the results: 
# Number-letter: switch 64, Repeat: 63
# Stroop: Congruent: 80 trials, Incongruent: 80 trials
# Spatial 2-back: Foil: 48 trials, Target: 112 trials
```

# Inferential statistics
### 1. Test for variance homogeneity
```{r}
# Ensure "Gruppe" is treated as a factor
Patients_vs_HC_imp$Gruppe <- as.factor(Patients_vs_HC_imp$Gruppe)

# Perform Levene's test for homogeneity of variance
levene_test_table <- levene_test_mult_cols(df_basis = Patients_vs_HC_imp, 
                                           cols = imp_columns, 
                                           grouping_variable = "Gruppe")
#pander(levene_test_table, style = "rmarkdown", fontsize = "tiny")
levene_test_table
```

### 2. Calculate Welch t-test
```{r}
t_test_table <- t_test_mult_cols(df_basis = Patients_vs_HC_imp, cols = imp_columns, 
                                 df_results_columns = ttest_results_columns, grouping_variable = "Gruppe")
#pander(t_test_table, style = "rmarkdown", fontsize = "tiny")
t_test_table
```

### 3. Prepare table for publication
```{r}

t_test_table_pub <- cbind(performance_measures = rownames(t_test_table), t_test_table)

# Combine "Mean" and "SD" columns and reduce dataframe
t_test_table_pub <- t_test_table_pub %>%
  mutate(
    `Patients Mean (SD)` = paste0(round(group_1_mean, 2), " (", round(group_1_sd, 2), ")"),
    `HC Mean (SD)` = paste0(round(group_0_mean, 2), " (", round(group_0_mean, 2), ")")
    ) %>%
  select(performance_measures, `Patients Mean (SD)`, `HC Mean (SD)`, 
         df, t_statistic, effect_size, p_value, p_value_adjusted)

# Rename performance measures using a named vector
rename_map <- c(
  'NumberLetter_BIS_Repeat' = "Number-Letter: Repeat",
  'NumberLetter_BIS_Switch' = "Number-Letter: Switch",
  'NumberLetter_BIS_Diff_Score' = "Number-Letter: Switch - Repeat",
  'Stroop_BIS_Congruent' = "Stroop: Congruent",
  'Stroop_BIS_Incongruent' = "Stroop: Incongruent",
  'Stroop_BIS_Diff_Score' = "Stroop: Incongruent - Congruent",
  'TwoBack_BIS_Foil' = "2-Back: Foil",
  'TwoBack_BIS_Target' = "2-Back: Target",
  'TwoBack_BIS_Total' = "2-Back: Total",
  'SSRT' = "Stop-Signal RT")
t_test_table_pub$performance_measures <- rename_map[t_test_table_pub$performance_measures]

# Change colnames
colnames(t_test_table_pub) <- c("Performance Measure", "Patients Mean (SD)", "HC Mean (SD)", "df", "t-Statistic", "Hedges` g", "p-Value", "adj. p-Value")
```

### 4. Create and save flextable
```{r}
ft_t_test <- flextable(t_test_table_pub)

# Set table properties
ft_t_test <- set_table_properties(ft_t_test, width = 1, layout = "autofit")

# Header in bold
ft_t_test <- bold(ft_t_test, bold = TRUE, part = "header")

# Alignments
ft_t_test <- align(ft_t_test, j = 1, align = "left", part = "all") # first column
ft_t_test <- align(ft_t_test, j = 2:ncol(t_test_table_pub), align = "center", part = "all") # rest

# Export flextable
save_as_docx(
  ft_t_test,
  path = file.path(results_path, "t-test_Patients_HCs.docx"), 
  pr_section = format_flextable_portrait)
```

# Plotting
### 1. Prepare data
```{r}
# Reshape data from wide to long format, including SSRT
Patients_vs_HC_long <- Patients_vs_HC_imp %>%
  pivot_longer(cols = all_of(imp_columns), # imp_columns now includes SSRT
               names_to = "Condition",
               values_to = "BIS_Score")

# Add new task variable to keep the headers intact for faceting
Patients_vs_HC_long <- Patients_vs_HC_long %>%
  mutate(Task = case_when(
    grepl("NumberLetter", Condition) ~ "Number-Letter Task",
    grepl("Stroop", Condition) ~ "Stroop Task",
    grepl("TwoBack", Condition) ~ "2-Back Task",
    Condition == "SSRT" ~ "Stop Signal Task"
  ))

# Remove prefixes from Condition labels, keeping Task intact
Patients_vs_HC_long <- Patients_vs_HC_long %>%
  mutate(Condition = gsub("NumberLetter_BIS_|Stroop_BIS_|TwoBack_BIS_", "", Condition)) %>%
  mutate(Condition = as.character(Condition)) %>%
  
  # Set the condition order for each task separately to avoid conflicts (cannot both be "Diff_Score")
  mutate(Condition = ifelse(Task == "NumberLetter" & Condition == "Diff_Score", "Diff_Score NL", Condition)) %>%
  mutate(Condition = ifelse(Task == "Stroop" & Condition == "Diff_Score", "Diff_Score Stroop", Condition)) %>%
  
  # Convert Condition to factor with ordering
  mutate(Condition = case_when(
    Task == "NumberLetter" ~ factor(Condition, levels = c("Repeat", "Switch", "Diff_Score NL")),
    Task == "Stroop" ~ factor(Condition, levels = c("Congruent", "Incongruent", "Diff_Score Stroop")),
    Task == "TwoBack" ~ factor(Condition, levels = c("Foil", "Target", "Total")),
    TRUE ~ as.factor(Condition)
  ))

# Convert "Gruppe" to factor
Patients_vs_HC_long$Gruppe <- factor(Patients_vs_HC_long$Gruppe, levels = c(0, 1), labels = c("Healthy Controls", "Patients"))
```

### 2. Violin Plot: Patients vs. Healthy Controls
```{r, fig.width = 16, fig.height = 9, warning = FALSE}
# Create violin plots with mean and rename conditions in the plot

violin_plots_patients_hc <- ggplot(Patients_vs_HC_long, aes(x = Condition, y = BIS_Score, fill = Gruppe)) +
  geom_violin(position = position_dodge(width = 0.8)) +
  stat_summary(fun.data = mean_sdl, geom = "pointrange", # mult = 2 (SD) by default
               position = position_dodge(width = 0.8)) + # width = 0.65, fatten = 2
  facet_wrap(~ Task, scales = "free") +
  labs(x = "",
       y = "Performance Score",
       fill = "Group") +
  scale_x_discrete(labels = function(x) {
    x[x == "Diff_Score"] <- "Difference \nScore"
    x[x == "Diff_Score Stroop"] <- "Difference \nScore"
    return(x)
  }) +
  scale_fill_manual(values = c("Healthy Controls" = "#D3D3D3", #colors 
                                 "Patients" = "#48CFCB")) +
  theme_bw() +
  theme(
    text = element_text(family = "Arial"), #Font
    axis.title = element_text(size = 24),
    axis.title.y = element_text(size = 24, margin = margin(r = 30)),
    axis.text = element_text(size = 18),
    legend.title = element_text(size = 24),
    legend.text = element_text(size = 18),
    strip.text = element_text(size = 20),
    panel.spacing.x = unit(3, "lines")
  )
print(violin_plots_patients_hc)
ggsave(file.path(results_path, "violin_plots_patients_hc.png"), 
       plot = violin_plots_patients_hc, width = 16, height = 9, dpi = 300)
ggsave(file.path(results_path, "violin_plots_patients_hc.svg"), 
       plot = violin_plots_patients_hc, width = 16, height = 9, dpi = 300)
```

### 3. Boxplot: Patients vs. Healthy Controls
```{r, warning = FALSE}
# Create boxplots
boxplots_Patients_HC <- ggplot(Patients_vs_HC_long, aes(x = Condition, y = BIS_Score, color = Gruppe)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(position = position_jitterdodge(),
              size = 0.7,
              shape = 21,
              fill = NA) +
  facet_wrap(~ Task, scales = "free") +
  #facet_grid(rows = vars(Task), scales = "free_x") +
  labs(title = "Boxplots of Task Condition by Group",
       x = "Condition",
       y = "BIS-Score",
       color = "Group")
print(boxplots_Patients_HC)
ggsave(file.path(results_path, "boxplots_patients_hc.png"), 
       plot = boxplots_Patients_HC, width = 16, height = 9, dpi = 300)
```
